### 2019-09-24 source/cadabra/checks/check-genGamma.tex

\documentclass[12pt]{cdblatex}

\begin{document}

% =================================================================================================
\section*{Compute $\Gamma^{a}{}_{bcd}$ directly}

Here we will compute one of generalised connections directly from the connection. That is, we will
compute
\begin{align}
   \Gamma^{a}{}_{bcd} = \Gamma^{a}{}_{(bc,d)} - 2 \Gamma^{a}{}_{p(c} \Gamma^{p}{}_{bd)}
\end{align}
given an explict expression for the RNC connection $\Gamma^{a}{}_{bc}$.

This code was written as a check for the {\tt genGamma.tex} code. I had a discrepency between my
newly created Cadabra v2.0 codes and my old Cadabra v1.0 codes (which were the basis of my lcb09-03 paper).
I found that my new codes agreed with this code and thus my old codes were wrong. I found the errors
in the old code (see the updated codes in v1.0/rnc-new/gen-gamma.cdbp, see also the file v1.0/rnc-new/NOTES.txt).

The head-on approach shown in this code works for this simple compuation but for the higher
generalised connections it's almost certainly going to be way too slow to be use. So this code is
only a check of the results from {\tt genGamma.tex} code. The good news is -- they agree.

\clearpage

\begin{cadabra}
   {a,b,c,d,e,f,g,h,i,j,k,l,m,n,o,p,q,r,s,t,u,v,w#}::Indices(position=independent).

   D{#}::Derivative.
   \nabla{#}::Derivative.
   \partial{#}::PartialDerivative.

   g_{a b}::Metric.
   g^{a b}::InverseMetric.
   g_{a}^{b}::KroneckerDelta.
   g^{a}_{b}::KroneckerDelta.
   \delta^{a}_{b}::KroneckerDelta.
   \delta_{a}^{b}::KroneckerDelta.

   R_{a b c d}::RiemannTensor.
   R_{a b c d}::Depends(\nabla{#}).

   x^{a}::Depends(\partial{#}).

   \Gamma^{a}_{b c}::Depends(\partial{#}).
   \Gamma^{a}_{b c}::TableauSymmetry(shape={2}, indices={1,2}).

   Q_{a b c d}::Weight(label=numR,value=2).
   Q_{a b c d e}::Weight(label=numR,value=3).
   Q_{a b c d e f}::Weight(label=numR,value=4).
   Q_{a b c d e f g}::Weight(label=numR,value=5).

   # note: keeping numbering as is (out of order) to ensure R appears before \nabla R etc.
   def product_sort (obj):
       substitute (obj,$ A^{a}                            -> A001^{a}               $)
       substitute (obj,$ x^{a}                            -> A002^{a}               $)
       substitute (obj,$ g^{a b}                          -> A003^{a b}             $)
       substitute (obj,$ \nabla_{e f g h}{R_{a b c d}}    -> A008_{a b c d e f g h} $)
       substitute (obj,$ \nabla_{e f g}{R_{a b c d}}      -> A007_{a b c d e f g}   $)
       substitute (obj,$ \nabla_{e f}{R_{a b c d}}        -> A006_{a b c d e f}     $)
       substitute (obj,$ \nabla_{e}{R_{a b c d}}          -> A005_{a b c d e}       $)
       substitute (obj,$ R_{a b c d}                      -> A004_{a b c d}         $)
       sort_product   (obj)
       rename_dummies (obj)
       substitute (obj,$ A001^{a}                  -> A^{a}                         $)
       substitute (obj,$ A002^{a}                  -> x^{a}                         $)
       substitute (obj,$ A003^{a b}                -> g^{a b}                       $)
       substitute (obj,$ A004_{a b c d}            -> R_{a b c d}                   $)
       substitute (obj,$ A005_{a b c d e}          -> \nabla_{e}{R_{a b c d}}       $)
       substitute (obj,$ A006_{a b c d e f}        -> \nabla_{e f}{R_{a b c d}}     $)
       substitute (obj,$ A007_{a b c d e f g}      -> \nabla_{e f g}{R_{a b c d}}   $)
       substitute (obj,$ A008_{a b c d e f g h}    -> \nabla_{e f g h}{R_{a b c d}} $)

       return obj

   def truncate (obj,n):

   # I would like to assign different weights to \nabla_{a}, \nabla_{a b}, \nabla_{a b c} etc. but no matter
   # what I do it appears that Cadabra assigns the same weight to all of these regardless of the number of subscripts.
   # It seems that the weight is assigned to the symbol \nabla alone. So I'm forced to use the following substitution trick.

       tmp := @(obj).

       substitute (tmp, $\nabla_{e f g}{R_{a b c d}} -> Q_{a b c d e f g}$)
       substitute (tmp, $\nabla_{e f}{R_{a b c d}} -> Q_{a b c d e f}$)
       substitute (tmp, $\nabla_{e}{R_{a b c d}} -> Q_{a b c d e}$)
       substitute (tmp, $R_{a b c d} -> Q_{a b c d}$)

       ans = Ex(0)

       for i in range (0,n+1):
          foo := @(tmp).
          bah = Ex("numR = " + str(i))
          keep_weight (foo, bah)
          ans = ans + foo

       substitute (ans, $Q_{a b c d e f g} -> \nabla_{e f g}{R_{a b c d}}$)
       substitute (ans, $Q_{a b c d e f} -> \nabla_{e f}{R_{a b c d}}$)
       substitute (ans, $Q_{a b c d e} -> \nabla_{e}{R_{a b c d}}$)
       substitute (ans, $Q_{a b c d} -> R_{a b c d}$)

       return ans

   def get_term (obj,n):

       bah := @(obj).
       distribute (bah)

       substitute (bah, $\nabla_{e f g}{R_{a b c d}} -> Q_{a b c d e f g}$)
       substitute (bah, $\nabla_{e f}{R_{a b c d}} -> Q_{a b c d e f}$)
       substitute (bah, $\nabla_{e}{R_{a b c d}} -> Q_{a b c d e}$)
       substitute (bah, $R_{a b c d} -> Q_{a b c d}$)

       foo = Ex("numR = " + str(n))
       keep_weight (bah, foo)

       substitute (bah, $Q_{a b c d e f g} -> \nabla_{e f g}{R_{a b c d}}$)
       substitute (bah, $Q_{a b c d e f} -> \nabla_{e f}{R_{a b c d}}$)
       substitute (bah, $Q_{a b c d e} -> \nabla_{e}{R_{a b c d}}$)
       substitute (bah, $Q_{a b c d} -> R_{a b c d}$)

       return bah

   def tidy (obj,number):
      bah  = Ex(str(number))
      tmp := @(bah) @(obj).
      distribute  (tmp)
      factor_out  (tmp,$A^{a?},x^{b?}$)
      ans := @(tmp) / @(bah).
      return ans

   import cdblib

   Gamma = cdblib.get ('Gamma','../connection.json')

   defGamma := \Gamma^{d}_{a b} -> @(Gamma).

   genGam := A^{d} A^{b} A^{c} (\partial_{d}{\Gamma^{a}_{b c}} - 2 \Gamma^{a}_{p c} \Gamma^{p}_{b d}).

   substitute     (genGam,defGamma)   # cdb (genGam.001,genGam)
   distribute     (genGam)
   unwrap         (genGam)            # cdb (genGam.002,genGam)
   distribute     (genGam)            # cdb (genGam.003,genGam)
   product_rule   (genGam)            # cdb (genGam.004,genGam)
   distribute     (genGam)            # cdb (genGam.005,genGam)
   substitute     (genGam,$\partial_{a}{x^{b}}->\delta_{a}^{b}$)   # cdb (genGam.006,genGam)
   eliminate_kronecker (genGam)       # cdb (genGam.007,genGam)
   genGam = truncate     (genGam,5)   # cdb (genGam.008,genGam)
   genGam = product_sort (genGam)     # cdb (genGam.009,genGam)
   rename_dummies (genGam)            # cdb (genGam.010,genGam)
   canonicalise   (genGam)            # cdb (genGam.011,genGam)

   # all done, now for some housekeeping

   term3 = get_term (genGam,3)   # cdb (term3.001,term3)
   term4 = get_term (genGam,4)   # cdb (term4.001,term4)
   term5 = get_term (genGam,5)   # cdb (term5.001,term5)

   term3 = tidy (term3,2)
   term4 = tidy (term4,120)
   term5 = tidy (term5,180)

   genGam := @(term3) + @(term4) + @(term5).

   scaledGamma := 360 @(genGam).  # cdb (scaledGamma.001,scaledGamma)

\end{cadabra}

\clearpage

\begin{dgroup*}
   \begin{dmath*} \cdb*{term3.001} \end{dmath*}
   \begin{dmath*} \cdb*{term4.001} \end{dmath*}
   \begin{dmath*} \cdb*{term5.001} \end{dmath*}
\end{dgroup*}

\clearpage

% =================================================================================================
\section*{Summary}

\begin{dgroup*}
   \begin{dmath*} 360 A^b A^c A^d \Gamma^{a}_{b c d} = \cdb{scaledGamma.001} \end{dmath*}
\end{dgroup*}

\end{document}