### 2019-09-24 source/cadabra/dGamma.tex

\documentclass[12pt]{cdblatex}

\begin{document}

% =================================================================================================
% create checkpoint file

\bgroup
\CdbSetup{action=hide}
\begin{cadabra}
   import cdblib
   checkpoint_file = 'tests/semantic/output/dGamma.json'
   cdblib.create (checkpoint_file)
   checkpoint = []
\end{cadabra}
\egroup

% =================================================================================================
\section*{Symmetrized partial derivatives of the connection}

Here we calculate the recursive sequences
\begin{align*}
(n+3)\Gamma^a{}_{d(b,c\uen)}
   &= (n+1)\left(R^a{}_{(bc\Dot d,\uen)}
    - \left(\Gamma^a{}_{f(c}\Gamma^f{}_{b{\Dot d}}\right){\vphantom{\Gamma}}_{,\uen)}\right)
\end{align*}
for $n=1,2,3,\cdots$. Note that the (extended) index $\uen$ contains $n$ normal indices.

The result will be expressions for the $\Gamma^a{}_{d(b,c\uen)}$ in terms of the Riemann tensor and its
partial derivatives.

% =================================================================================================
\section*{Stage 1: Compute symmetrised derivatives}

In the first stage we simply apply the above recursive equation using a simple trick to impose
the symmetries. Start with the original equation and dot out the symmetric indices with $A^a$ then
factor out the partial derivatives. This leads to
\begin{equation}
(n+3)\Gamma^a{}_{db,c\uen} A^b A^c A^{\cdot\uen}
    = (n+1)\left(R^a{}_{bcd}
    - \Gamma^a{}_{fc}\Gamma^f{}_{bd}\right)_{,\uen} A^b A^c A^{\cdot\uen}
\end{equation}
Thus we also have (for the next iteration)
\begin{equation}
(n+4)\Gamma^a{}_{db,c\uenp} A^b A^c A^{\cdot\uenp}
    = (n+2)\left(R^a{}_{bcd}
    - \Gamma^a{}_{fc}\Gamma^f{}_{bd}\right)_{,\uenp} A^b A^c A^{\cdot\uenp}
\end{equation}
The $A^a$ can be freely chosen so choose $A^a$ to be a constant (i.e., zero derivative). Now
define $P_n$ by
\begin{equation}
   P_n = \Gamma^a{}_{db,c\uen} A^b A^c A^{\cdot\uen}
\end{equation}
then the above pair of equations can be combinded to give
\begin{equation}
   P_{n+1} = \frac{(n+2)(n+3)}{(n+4)(n+1)} A^f\partial_f\left(P_n\right)
\end{equation}
This is a very easy equation to compute as it just requires successive rounds of differentiation.

The first term in the sequence is $P_0$ given by
\begin{equation}
   P_0 = \cdb{dGamma01.101}
\end{equation}

The first few results are

\begin{dgroup*}
   \begin{dmath*} P_0 \hiderel{=} A^{b} A^{c} \Gamma^a{}_{d(b,c)} = \cdb{dGamma01.101} \end{dmath*}
   \begin{dmath*} P_1 \hiderel{=} A^{b} A^{c} A^{e} \Gamma^a{}_{d(b,ce)} = \cdb{dGamma02.105} \end{dmath*}
   \begin{dmath*} P_2 \hiderel{=} A^{b} A^{c} A^{e} A^{f} \Gamma^a{}_{d(b,cef)} = \cdb{dGamma03.105} \end{dmath*}
\end{dgroup*}

% =================================================================================================
\section*{Stage 2: Impose Riemann normal coordinates}

Here we impose the RNC conition by setting the $\Gamma^{a}{}_{bc}$ to zero (but not their derivatives).

\begin{dgroup*}
   \begin{dmath*} A^{b} A^{c} \Gamma^a{}_{d(b,c)} = \cdb{dGamma01.201} \end{dmath*}
   \begin{dmath*} A^{b} A^{c} A^{e} \Gamma^a{}_{d(b,ce)} = \cdb{dGamma02.202} \end{dmath*}
   \begin{dmath*} A^{b} A^{c} A^{e} A^{f} \Gamma^a{}_{d(b,cef)} = \cdb{dGamma03.203} \end{dmath*}
\end{dgroup*}

% =================================================================================================
\section*{Stage 3: Replace partial derivatives $\Gamma$ with partial derivatives $R$}

The key point to note is that the partial derivatives of $\Gamma$ on the right hand side are all
symmetrized in exactly the same manner as the partial derivatives on the left hand side. Thus
results from the lower order equations can be fed into the later equations to completely eliminate
the partial derivatives of $\Gamma$.

\begin{dgroup*}
   \begin{dmath*} A^{b} A^{c} \Gamma^a{}_{d(b,c)} = \cdb{dGamma01.201} \end{dmath*}
   \begin{dmath*} A^{b} A^{c} A^{e} \Gamma^a{}_{d(b,ce)} = \cdb{dGamma02.202} \end{dmath*}
   \begin{dmath*} A^{b} A^{c} A^{e} A^{f} \Gamma^a{}_{d(b,cef)} = \cdb{dGamma03.402} \end{dmath*}
\end{dgroup*}

% =================================================================================================
\section*{Stage 4: Reformatting}

This is just simple reformatting.

\begin{dgroup*}
   \begin{dmath*}   3 A^b A^c\Gamma^a{}_{d(b,c)} = \cdb{scaled1.002} \end{dmath*}
   \begin{dmath*}   6 A^b A^c A^e \Gamma^a{}_{d(b,ce)} = \cdb{scaled2.002} \end{dmath*}
   \begin{dmath*}  15 A^b A^c A^e A^f \Gamma^a{}_{d(b,cef)} = \cdb{scaled3.002} \end{dmath*}
\end{dgroup*}

\clearpage

% =================================================================================================
\section*{Stage 1: Compute symmetrised derivatives}

\begin{cadabra}
   {a,b,c,d,e,f,g,h,i,j,k,l,m,n,o,p,q,r,s,t,u,v,w#}::Indices(position=independent).

   \nabla{#}::Derivative.
   \partial{#}::PartialDerivative.

   g_{a b}::Metric.
   g^{a b}::InverseMetric.
   g_{a}^{b}::KroneckerDelta.
   g^{a}_{b}::KroneckerDelta.

   R_{a b c d}::RiemannTensor.
   R^{a}_{b c d}::RiemannTensor.
   R_{a b c}^{d}::RiemannTensor.

   \Gamma^{a}_{b c}::TableauSymmetry(shape={2}, indices={1,2}).

   g_{a b}::Depends(\partial{#}).
   R_{a b c d}::Depends(\partial{#}).
   R^{a}_{b c d}::Depends(\partial{#}).
   \Gamma^{a}_{b c}::Depends(\partial{#}).

   # symmetrized partial derivatives of \Gamma

   dGamma01:= (1/3) A^{b} A^{c} ( R^{a}_{b c d} - \Gamma^{a}_{c e}\Gamma^{e}_{b d} ).
                                                        # cdb (dGamma01.101,dGamma01)

   dGamma02:= (6/4) A^{a}\partial_{a}{ @(dGamma01) }.   # cdb (dGamma02.101,dGamma02)
   distribute   (dGamma02)                              # cdb (dGamma02.102,dGamma02)
   product_rule (dGamma02)                              # cdb (dGamma02.103,dGamma02)
   unwrap       (dGamma02)                              # cdb (dGamma02.104,dGamma02)
   distribute   (dGamma02)                              # cdb (dGamma02.105,dGamma02)

   dGamma03:= (12/10) A^{a}\partial_{a}{ @(dGamma02) }. # cdb (dGamma03.101,dGamma03)
   distribute   (dGamma03)                              # cdb (dGamma03.102,dGamma03)
   product_rule (dGamma03)                              # cdb (dGamma03.103,dGamma03)
   unwrap       (dGamma03)                              # cdb (dGamma03.104,dGamma03)
   distribute   (dGamma03)                              # cdb (dGamma03.105,dGamma03)

   dGamma04:= (20/18) A^{a}\partial_{a}{ @(dGamma03) }. # cdb (dGamma04.101,dGamma04)
   distribute   (dGamma04)                              # cdb (dGamma04.102,dGamma04)
   product_rule (dGamma04)                              # cdb (dGamma04.103,dGamma04)
   unwrap       (dGamma04)                              # cdb (dGamma04.104,dGamma04)
   distribute   (dGamma04)                              # cdb (dGamma04.105,dGamma04)

   dGamma05:= (30/28) A^{a}\partial_{a}{ @(dGamma04) }. # cdb (dGamma05.101,dGamma05)
   distribute   (dGamma05)                              # cdb (dGamma05.102,dGamma05)
   product_rule (dGamma05)                              # cdb (dGamma05.103,dGamma05)
   unwrap       (dGamma05)                              # cdb (dGamma05.104,dGamma05)
   distribute   (dGamma05)                              # cdb (dGamma05.105,dGamma05)

\end{cadabra}

\clearpage

\begin{dgroup*}
   \begin{dmath*} \cdb*{dGamma01.101} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{dGamma02.101} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma02.102} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma02.103} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma02.104} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma02.105} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{dGamma03.101} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma03.102} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma03.103} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma03.104} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma03.105} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{dGamma04.101} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma04.102} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma04.103} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma04.104} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma04.105} \end{dmath*}
\end{dgroup*}

% too long for pdflatex

% \begin{dgroup*}
%    \begin{dmath*} \cdb*{dGamma05.101} \end{dmath*}
%    \begin{dmath*} \cdb*{dGamma05.102} \end{dmath*}
%    \begin{dmath*} \cdb*{dGamma05.103} \end{dmath*}
%    \begin{dmath*} \cdb*{dGamma05.104} \end{dmath*}
%    \begin{dmath*} \cdb*{dGamma05.105} \end{dmath*}
% \end{dgroup*}

\clearpage

% =================================================================================================
\section*{Stage 2: Impose Riemann normal coordinates}

\begin{cadabra}
   def impose_rnc (obj):
       # hide the derivatives of Gamma
       substitute (obj,$\partial_{d}{\Gamma^{a}_{b c}} -> zzz_{d}^{a}_{b c}$,repeat=True)
       substitute (obj,$\partial_{d e}{\Gamma^{a}_{b c}} -> zzz_{d e}^{a}_{b c}$,repeat=True)
       substitute (obj,$\partial_{d e f}{\Gamma^{a}_{b c}} -> zzz_{d e f}^{a}_{b c}$,repeat=True)
       substitute (obj,$\partial_{d e f g}{\Gamma^{a}_{b c}} -> zzz_{d e f g}^{a}_{b c}$,repeat=True)
       substitute (obj,$\partial_{d e f g h}{\Gamma^{a}_{b c}} -> zzz_{d e f g h}^{a}_{b c}$,repeat=True)
       # set Gamma to zero
       substitute (obj,$\Gamma^{a}_{b c} -> 0$,repeat=True)
       # recover the derivatives Gamma
       substitute (obj,$zzz_{d}^{a}_{b c} -> \partial_{d}{\Gamma^{a}_{b c}}$,repeat=True)
       substitute (obj,$zzz_{d e}^{a}_{b c} -> \partial_{d e}{\Gamma^{a}_{b c}}$,repeat=True)
       substitute (obj,$zzz_{d e f}^{a}_{b c} -> \partial_{d e f}{\Gamma^{a}_{b c}}$,repeat=True)
       substitute (obj,$zzz_{d e f g}^{a}_{b c} -> \partial_{d e f g}{\Gamma^{a}_{b c}}$,repeat=True)
       substitute (obj,$zzz_{d e f g h}^{a}_{b c} -> \partial_{d e f g h}{\Gamma^{a}_{b c}}$,repeat=True)
       return obj

   # switch to RNC

   dGamma01 = impose_rnc (dGamma01)   # cdb (dGamma01.201,dGamma01)
   dGamma02 = impose_rnc (dGamma02)   # cdb (dGamma02.202,dGamma02)
   dGamma03 = impose_rnc (dGamma03)   # cdb (dGamma03.203,dGamma03)
   dGamma04 = impose_rnc (dGamma04)   # cdb (dGamma04.204,dGamma04)
   dGamma05 = impose_rnc (dGamma05)   # cdb (dGamma05.205,dGamma05)

\end{cadabra}

\begin{dgroup*}
   \begin{dmath*} \cdb*{dGamma01.201} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma02.202} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma03.203} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma04.204} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma05.205} \end{dmath*}
\end{dgroup*}

\clearpage

% =================================================================================================
\section*{Stage 3: Replace partial derivatives $\Gamma$ with partial derivatives $R$}

\begin{cadabra}
   # use lower equations to eliminate partial derivs of Gamma from rhs

   # this produces experssions for the partial derivs of the Gamma's in terms of the Rabcd and its partial derivs

   substitute (dGamma03,$A^{c}A^{b}\partial_{c}{\Gamma^{a}_{b d}} -> @(dGamma01)$,repeat=True)                 # cdb(dGamma03.301,dGamma03)
   substitute (dGamma03,$A^{c}A^{b}\partial_{c}{\Gamma^{a}_{d b}} -> @(dGamma01)$,repeat=True)                 # cdb(dGamma03.302,dGamma03)
   distribute (dGamma03)                                                                                       # cdb(dGamma03.303,dGamma03)

   substitute (dGamma04,$A^{c}A^{b}A^{e}\partial_{c e}{\Gamma^{a}_{d b}} -> @(dGamma02)$,repeat=True)          # cdb(dGamma04.301,dGamma04)
   substitute (dGamma04,$A^{c}A^{b}A^{e}\partial_{c e}{\Gamma^{a}_{b d}} -> @(dGamma02)$,repeat=True)          # cdb(dGamma04.302,dGamma04)
   substitute (dGamma04,$A^{c}A^{b}\partial_{c}{\Gamma^{a}_{b d}} -> @(dGamma01)$,repeat=True)                 # cdb(dGamma04.303,dGamma04)
   substitute (dGamma04,$A^{c}A^{b}\partial_{c}{\Gamma^{a}_{d b}} -> @(dGamma01)$,repeat=True)                 # cdb(dGamma04.304,dGamma04)
   distribute (dGamma04)                                                                                       # cdb(dGamma04.305,dGamma04)

   substitute (dGamma05,$A^{c}A^{b}A^{e}A^{f}\partial_{c e f}{\Gamma^{a}_{d b}} -> @(dGamma03)$,repeat=True)   # cdb(dGamma05.301,dGamma05)
   substitute (dGamma05,$A^{c}A^{b}A^{e}A^{f}\partial_{c e f}{\Gamma^{a}_{b d}} -> @(dGamma03)$,repeat=True)   # cdb(dGamma05.302,dGamma05)
   substitute (dGamma05,$A^{c}A^{b}A^{e}\partial_{c e}{\Gamma^{a}_{d b}} -> @(dGamma02)$,repeat=True)          # cdb(dGamma05.303,dGamma05)
   substitute (dGamma05,$A^{c}A^{b}A^{e}\partial_{c e}{\Gamma^{a}_{b d}} -> @(dGamma02)$,repeat=True)          # cdb(dGamma05.304,dGamma05)
   substitute (dGamma05,$A^{c}A^{b}\partial_{c}{\Gamma^{a}_{b d}} -> @(dGamma01)$,repeat=True)                 # cdb(dGamma05.305,dGamma05)
   substitute (dGamma05,$A^{c}A^{b}\partial_{c}{\Gamma^{a}_{d b}} -> @(dGamma01)$,repeat=True)                 # cdb(dGamma05.306,dGamma05)
   distribute (dGamma05)                                                                                       # cdb(dGamma05.307,dGamma05)

\end{cadabra}

\clearpage

\begin{dgroup*}
   \begin{dmath*} \cdb*{dGamma03.301} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma03.302} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma03.303} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{dGamma04.301} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma04.302} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma04.303} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma04.304} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma04.305} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{dGamma05.301} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma05.302} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma05.303} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma05.304} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma05.305} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma05.306} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma05.307} \end{dmath*}
\end{dgroup*}

\clearpage

\begin{cadabra}
   # note:
   # canonicalise must not be used here because it may make changes like
   #    R^{a}_{b c d} -> - R_{b}^{a}_{c d}
   # these changes can not be applied inside a \partial, must defer use
   # of canocialise until we have \nabla acting on curvatures

   sort_product   (dGamma03) # cdb(dGamma03.401,dGamma03)
   rename_dummies (dGamma03) # cdb(dGamma03.402,dGamma03)
   # canonicalise   (dGamma03) # cdb(dGamma03.403,dGamma03)

   sort_product   (dGamma04) # cdb(dGamma04.401,dGamma04)
   rename_dummies (dGamma04) # cdb(dGamma04.402,dGamma04)
   # canonicalise   (dGamma04) # cdb(dGamma04.403,dGamma04)

   sort_product   (dGamma05) # cdb(dGamma05.401,dGamma05)
   rename_dummies (dGamma05) # cdb(dGamma05.402,dGamma05)
   # canonicalise   (dGamma05) # cdb(dGamma05.403,dGamma05)

\end{cadabra}

\clearpage

\begin{dgroup*}
   \begin{dmath*} \cdb*{dGamma03.401} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma03.402} \end{dmath*}
   % \begin{dmath*} \cdb*{dGamma03.403} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{dGamma04.401} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma04.402} \end{dmath*}
   % \begin{dmath*} \cdb*{dGamma04.403} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{dGamma05.401} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma05.402} \end{dmath*}
   % \begin{dmath*} \cdb*{dGamma05.403} \end{dmath*}
\end{dgroup*}

\clearpage

\begin{cadabra}
   import cdblib

   cdblib.create ('dGamma.json')

   cdblib.put ('dGamma01',dGamma01,'dGamma.json')
   cdblib.put ('dGamma02',dGamma02,'dGamma.json')
   cdblib.put ('dGamma03',dGamma03,'dGamma.json')
   cdblib.put ('dGamma04',dGamma04,'dGamma.json')
   cdblib.put ('dGamma05',dGamma05,'dGamma.json')

\end{cadabra}

\clearpage

% =================================================================================================
\section*{Stage 4: Reformatting}

\begin{cadabra}
   # note: keeping numbering as is (out of order) to ensure R appears before \nabla R etc.
   def product_sort (obj):
       substitute (obj,$ A^{a}                              -> A001^{a}                 $)
       substitute (obj,$ x^{a}                              -> A002^{a}                 $)
       substitute (obj,$ g^{a b}                            -> A003^{a b}               $)
       substitute (obj,$ \partial_{e f g h}{R^{a}_{b c d}}  -> A008^{a}_{b c d e f g h} $)
       substitute (obj,$ \partial_{e f g}{R^{a}_{b c d}}    -> A007^{a}_{b c d e f g}   $)
       substitute (obj,$ \partial_{e f}{R^{a}_{b c d}}      -> A006^{a}_{b c d e f}     $)
       substitute (obj,$ \partial_{e}{R^{a}_{b c d}}        -> A005^{a}_{b c d e}       $)
       substitute (obj,$ R^{a}_{b c d}                      -> A004^{a}_{b c d}         $)
       sort_product   (obj)
       rename_dummies (obj)
       substitute (obj,$ A001^{a}                  -> A^{a}                             $)
       substitute (obj,$ A002^{a}                  -> x^{a}                             $)
       substitute (obj,$ A003^{a b}                -> g^{a b}                           $)
       substitute (obj,$ A004^{a}_{b c d}          -> R^{a}_{b c d}                     $)
       substitute (obj,$ A005^{a}_{b c d e}        -> \partial_{e}{R^{a}_{b c d}}       $)
       substitute (obj,$ A006^{a}_{b c d e f}      -> \partial_{e f}{R^{a}_{b c d}}     $)
       substitute (obj,$ A007^{a}_{b c d e f g}    -> \partial_{e f g}{R^{a}_{b c d}}   $)
       substitute (obj,$ A008^{a}_{b c d e f g h}  -> \partial_{e f g h}{R^{a}_{b c d}} $)

       return obj

   def reformat (obj,scale):
       bah  = Ex(str(scale))
       tmp := @(bah) @(obj).
       distribute     (tmp)
       tmp = product_sort (tmp)
       rename_dummies (tmp)
       factor_out     (tmp,$A^{a?}$)
       return tmp

   def get_term (obj,n):

       A^{a}::Weight(label=numA).

       foo := @(obj).
       bah  = Ex("numA = " + str(n))
       distribute  (foo)
       keep_weight (foo, bah)

       return foo

   Gterm01 := @(dGamma01).
   Gterm02 := @(dGamma02).
   Gterm03 := @(dGamma03).
   Gterm04 := @(dGamma04).
   Gterm05 := @(dGamma05).

   scaled1 = reformat (Gterm01,   3)   # cdb (scaled1.002,scaled1)
   scaled2 = reformat (Gterm02,   6)   # cdb (scaled2.002,scaled2)
   scaled3 = reformat (Gterm03,  15)   # cdb (scaled3.002,scaled3)
   scaled4 = reformat (Gterm04,   9)   # cdb (scaled4.002,scaled4)
   scaled5 = reformat (Gterm05, 252)   # cdb (scaled5.002,scaled5)

\end{cadabra}

\clearpage

% =================================================================================================
\section*{Symmetrised partial derivatives of the connection}

\begin{dgroup*}
   \begin{dmath*}   3 A^b A^c\Gamma^a{}_{d(b,c)} = \cdb{scaled1.002} \end{dmath*}
   \begin{dmath*}   6 A^b A^c A^e \Gamma^a{}_{d(b,ce)} = \cdb{scaled2.002} \end{dmath*}
   \begin{dmath*}  15 A^b A^c A^e A^f \Gamma^a{}_{d(b,cef)} = \cdb{scaled3.002} \end{dmath*}
   \begin{dmath*}   9 A^b A^c A^e A^f A^g \Gamma^a{}_{d(b,cefg)} = \cdb{scaled4.002} \end{dmath*}
   \begin{dmath*} 252 A^b A^c A^e A^f A^g A^h\Gamma^a{}_{d(b,cefgh)} = \cdb{scaled5.002} \end{dmath*}
\end{dgroup*}

\clearpage

% =================================================================================================
% export selected objects, these will later be imported into a library
% these are the objects that will appear in the paper

\begin{cadabra}
   substitute (scaled1,$A^{a}->1$)
   substitute (scaled2,$A^{a}->1$)
   substitute (scaled3,$A^{a}->1$)
   substitute (scaled4,$A^{a}->1$)
   substitute (scaled5,$A^{a}->1$)

   cdblib.create ('dGamma.export')

   # 6th order dGamma, scaled
   cdblib.put ('dGamma61scaled',scaled1,'dGamma.export')
   cdblib.put ('dGamma62scaled',scaled2,'dGamma.export')
   cdblib.put ('dGamma63scaled',scaled3,'dGamma.export')
   cdblib.put ('dGamma64scaled',scaled4,'dGamma.export')
   cdblib.put ('dGamma65scaled',scaled5,'dGamma.export')

   checkpoint.append (scaled1)
   checkpoint.append (scaled2)
   checkpoint.append (scaled3)
   checkpoint.append (scaled4)
   checkpoint.append (scaled5)

\end{cadabra}

% =================================================================================================
% export checkpoints in json format

\bgroup
\CdbSetup{action=hide}
\begin{cadabra}
   for i in range( len(checkpoint) ):
      cdblib.put ('check{:03d}'.format(i),checkpoint[i],checkpoint_file)
\end{cadabra}
\egroup

\end{document}