### 2019-09-24 source/cadabra/metric-inv.tex

\documentclass[12pt]{cdblatex}

\begin{document}

% =================================================================================================
% create checkpoint file

\bgroup
\CdbSetup{action=hide}
\begin{cadabra}
   import cdblib
   checkpoint_file = 'tests/semantic/output/metric-inv.json'
   cdblib.create (checkpoint_file)
   checkpoint = []
\end{cadabra}
\egroup

% =================================================================================================
\section*{The inverse metric tensor in Riemann normal coordinates}

Here we calculate the Riemann normal expansion of the inverse metric, $g^{ab}$, by developing
the recursive sequences
\begin{align}
\label{eq:pdgab}
g^{ab}{}_{,d\ue} &= - \left(g^{cb}\Gamma^{a}{}_{c(d}\right){}_{,\ue)}
                    - \left(g^{ac}\Gamma^{b}{}_{c(d}\right){}_{,\ue)}\\[10pt]
\label{eq:pdGamma}
(n+3)\Gamma^a{}_{d(b,c\ue)} &= (n+1)\left(R^a{}_{(bc\Dot d,\ue)} - \left(\Gamma^a{}_{f(c}\Gamma^f{}_{b{\Dot d}}\right){}_{,\ue)}\right)
\end{align}
for $n=1,2,3,\cdots$. Note in these equations that the (extended) index $\ue$ contains $n$
normal indices.

We then construct a Taylor series for the metric using
\begin{dmath*}[spread=5pt]
g^{ab}(x) = g^{ab} + g^{ab}{}_{,c}x^c + \frac{1}{2!} g^{ab}{}_{,cd}x^cx^d + \frac{1}{3!} g^{ab}{}_{,cde}x^cx^dx^e + \cdots
          = g^{ab} + \sum_{n=1}^\infty\> \frac{1}{n!}\>g^{ab}{}_{,\uc}\>x^{.\uc}
\end{dmath*}

% =================================================================================================
\section*{Stage 1: Symmetrised partial derivatives of $g^{ab}$}

In this stage, equation (\ref{eq:pdgab}) is used to express the symmetrised partial derivatives
of the metric in terms of the symmetrised partial derivatives of the connection.

\begin{dgroup*}
   \begin{dmath*} g^{ab}{}_{,c} A^{c} = \cdb{term1.200} \end{dmath*}
   \begin{dmath*} g^{ab}{}_{,cd} A^{c} A^{d} = \cdb{term2.200} \end{dmath*}
   \begin{dmath*} g^{ab}{}_{,cde} A^{c} A^{d} A^{e} = \cdb{term3.200} \end{dmath*}
\end{dgroup*}

% =================================================================================================
\section*{Stage 2: Replace derivatives of $\Gamma$ with partial derivs of $R$}

Now we use the results from {\verb|dGamma|} to replace derivatives of $\Gamma$ with
partial derivatives of $R$. These were computed in {\verb|dGamma|} using equation
(\ref{eq:pdGamma}) above.

\begin{dgroup*}
   \begin{dmath*} g^{ab}{}_{,c} A^{c} = \cdb{term1.200} \end{dmath*}
   \begin{dmath*} g^{ab}{}_{,cd} A^{c} A^{d} = \cdb{term2.303} \end{dmath*}
   \begin{dmath*} g^{ab}{}_{,cde} A^{c} A^{d} A^{e} = \cdb{term3.305} \end{dmath*}
\end{dgroup*}

% =================================================================================================
\section*{Stage 3: Replace partial derivs of $R$ with covariant derivs of $R$}

Next we use the results from {\verb|dRabcd|} to replace the partial derivatives of $R$ with
covariant deriavtives.

\begin{dgroup*}
   \begin{dmath*} g^{ab}{}_{,c} A^{c} = \cdb{term1.404} \end{dmath*}
   \begin{dmath*} g^{ab}{}_{,cd} A^{c} A^{d} = \cdb{term2.404} \end{dmath*}
   \begin{dmath*} g^{ab}{}_{,cde} A^{c} A^{d} A^{e} = \cdb{term3.403} \end{dmath*}
\end{dgroup*}

% =================================================================================================
\section*{Stage 4: Build the Taylor series for $g_{ab}$, reformatting and output}

Each of the above expressions constitutues one term in the Taylor series for the metric.
We also make the trivial change $A\rightarrow x$. Then we do some trivial reformatting.

\begin{align*}
   g_{ab}(x) &=   g^{ab}
                + g^{ab}{}_{,c} x^c
                + \frac{1}{2!} g^{ab}{}_{,cd} x^c x^d
                + \frac{1}{3!} g^{ab}{}_{,cde} x^c x^d x^e +  \BigO{\eps^4}\\
             &= \cdb{metric4.501} + \BigO{\eps^4}
\end{align*}

\clearpage

% =================================================================================================
\section*{Shared properties}

\begin{cadabra}
   import time

   def flatten_Rabcd (obj):
       substitute (obj,$R^{a}_{b c d}   -> g^{a e} R_{e b c d}$)
       substitute (obj,$R_{a}^{b}_{c d} -> g^{b e} R_{a e c d}$)
       substitute (obj,$R_{a b}^{c}_{b} -> g^{c e} R_{a b e d}$)
       substitute (obj,$R_{a b c}^{d}   -> g^{d e} R_{a b c e}$)
       unwrap     (obj)
       sort_product   (obj)
       rename_dummies (obj)
       return obj

   def impose_rnc (obj):
       # hide the derivatives of Gamma
       substitute (obj,$\partial_{d}{\Gamma^{a}_{b c}} -> zzz_{d}^{a}_{b c}$,repeat=True)
       substitute (obj,$\partial_{d e}{\Gamma^{a}_{b c}} -> zzz_{d e}^{a}_{b c}$,repeat=True)
       substitute (obj,$\partial_{d e f}{\Gamma^{a}_{b c}} -> zzz_{d e f}^{a}_{b c}$,repeat=True)
       substitute (obj,$\partial_{d e f g}{\Gamma^{a}_{b c}} -> zzz_{d e f g}^{a}_{b c}$,repeat=True)
       substitute (obj,$\partial_{d e f g h}{\Gamma^{a}_{b c}} -> zzz_{d e f g h}^{a}_{b c}$,repeat=True)
       # set Gamma to zero
       substitute (obj,$\Gamma^{a}_{b c} -> 0$,repeat=True)
       # recover the derivatives Gamma
       substitute (obj,$zzz_{d}^{a}_{b c} -> \partial_{d}{\Gamma^{a}_{b c}}$,repeat=True)
       substitute (obj,$zzz_{d e}^{a}_{b c} -> \partial_{d e}{\Gamma^{a}_{b c}}$,repeat=True)
       substitute (obj,$zzz_{d e f}^{a}_{b c} -> \partial_{d e f}{\Gamma^{a}_{b c}}$,repeat=True)
       substitute (obj,$zzz_{d e f g}^{a}_{b c} -> \partial_{d e f g}{\Gamma^{a}_{b c}}$,repeat=True)
       substitute (obj,$zzz_{d e f g h}^{a}_{b c} -> \partial_{d e f g h}{\Gamma^{a}_{b c}}$,repeat=True)
       return obj

   def get_xterm (obj,n):

       x^{a}::Weight(label=numx).

       foo := @(obj).
       bah  = Ex("numx = " + str(n))
       keep_weight (foo,bah)

       return foo

   # note: keeping numbering as is (out of order) to ensure R appears before \nabla R etc.
   def product_sort (obj):
       substitute (obj,$ A^{a}                             -> A001^{a}                  $)
       substitute (obj,$ x^{a}                             -> A002^{a}                  $)
       substitute (obj,$ g_{a b}                           -> A003_{a b}                $)
       substitute (obj,$ g^{a b}                           -> A004^{a b}                $)
       substitute (obj,$ \nabla_{e f g h}{R_{a b c d}}     -> A010_{a b c d e f g h}    $)
       substitute (obj,$ \nabla_{e f g}{R_{a b c d}}       -> A009_{a b c d e f g}      $)
       substitute (obj,$ \nabla_{e f}{R_{a b c d}}         -> A008_{a b c d e f}        $)
       substitute (obj,$ \nabla_{e}{R_{a b c d}}           -> A007_{a b c d e}          $)
       substitute (obj,$ \partial_{e f g h}{R_{a b c d}}   -> A014_{a b c d e f g h}    $)
       substitute (obj,$ \partial_{e f g}{R_{a b c d}}     -> A013_{a b c d e f g}      $)
       substitute (obj,$ \partial_{e f}{R_{a b c d}}       -> A012_{a b c d e f}        $)
       substitute (obj,$ \partial_{e}{R_{a b c d}}         -> A011_{a b c d e}          $)
       substitute (obj,$ \partial_{e f g h}{R^{a}_{b c d}} -> A018^{a}_{b c d e f g h}  $)
       substitute (obj,$ \partial_{e f g}{R^{a}_{b c d}}   -> A017^{a}_{b c d e f g}    $)
       substitute (obj,$ \partial_{e f}{R^{a}_{b c d}}     -> A016^{a}_{b c d e f}      $)
       substitute (obj,$ \partial_{e}{R^{a}_{b c d}}       -> A015^{a}_{b c d e}        $)
       substitute (obj,$ R_{a b c d}                       -> A005_{a b c d}            $)
       substitute (obj,$ R^{a}_{b c d}                     -> A006^{a}_{b c d}          $)
       sort_product   (obj)
       rename_dummies (obj)
       substitute (obj,$ A001^{a}                  -> A^{a}                             $)
       substitute (obj,$ A002^{a}                  -> x^{a}                             $)
       substitute (obj,$ A003_{a b}                -> g_{a b}                           $)
       substitute (obj,$ A004^{a b}                -> g^{a b}                           $)
       substitute (obj,$ A005_{a b c d}            -> R_{a b c d}                       $)
       substitute (obj,$ A006^{a}_{b c d}          -> R^{a}_{b c d}                     $)
       substitute (obj,$ A007_{a b c d e}          -> \nabla_{e}{R_{a b c d}}           $)
       substitute (obj,$ A008_{a b c d e f}        -> \nabla_{e f}{R_{a b c d}}         $)
       substitute (obj,$ A009_{a b c d e f g}      -> \nabla_{e f g}{R_{a b c d}}       $)
       substitute (obj,$ A010_{a b c d e f g h}    -> \nabla_{e f g h}{R_{a b c d}}     $)
       substitute (obj,$ A011_{a b c d e}          -> \partial_{e}{R_{a b c d}}         $)
       substitute (obj,$ A012_{a b c d e f}        -> \partial_{e f}{R_{a b c d}}       $)
       substitute (obj,$ A013_{a b c d e f g}      -> \partial_{e f g}{R_{a b c d}}     $)
       substitute (obj,$ A014_{a b c d e f g h}    -> \partial_{e f g h}{R_{a b c d}}   $)
       substitute (obj,$ A015^{a}_{b c d e}        -> \partial_{e}{R^{a}_{b c d}}       $)
       substitute (obj,$ A016^{a}_{b c d e f}      -> \partial_{e f}{R^{a}_{b c d}}     $)
       substitute (obj,$ A017^{a}_{b c d e f g}    -> \partial_{e f g}{R^{a}_{b c d}}   $)
       substitute (obj,$ A018^{a}_{b c d e f g h}  -> \partial_{e f g h}{R^{a}_{b c d}} $)

       return obj

   def reformat_xterm (obj,scale):
      foo  = Ex(str(scale))
      bah := @(foo) @(obj).
      distribute     (bah)
      bah = product_sort (bah)
      rename_dummies (bah)
      canonicalise   (bah)
      factor_out     (bah,$x^{a?}$)
      ans := @(bah) / @(foo).
      return ans

   def rescale_xterm (obj,scale):
      foo  = Ex(str(scale))
      bah := @(foo) @(obj).
      distribute  (bah)
      factor_out  (bah,$x^{a?}$)
      return bah

   def add_tags (obj,tag):
      n = 0
      ans = Ex('0')
      for i in obj.top().terms():
         foo = obj[i]
         bah = Ex(tag+'_{'+str(n)+'}')
         ans := @(ans) + @(bah) @(foo).
         n = n + 1
      return ans

   def clear_tags (obj,tag):
      ans := @(obj).
      foo  = Ex(tag+'_{a?} -> 1')
      substitute (ans,foo)
      return ans

   {a,b,c,d,e,f,g,h,i,j,k,l,m,n,o,p,q,r,s,t,u,v,w#}::Indices(position=independent).

   \nabla{#}::Derivative.
   \partial{#}::PartialDerivative.

   g_{a b}::Metric.
   g^{a b}::InverseMetric.
   g_{a}^{b}::KroneckerDelta.
   g^{a}_{b}::KroneckerDelta.

   R_{a b c d}::RiemannTensor.
   R^{a}_{b c d}::RiemannTensor.
   R_{a b c}^{d}::RiemannTensor.

   \Gamma^{a}_{b c}::TableauSymmetry(shape={2}, indices={1,2}).

   g_{a b}::Depends(\partial{#}).
   R_{a b c d}::Depends(\partial{#}).
   R^{a}_{b c d}::Depends(\partial{#}).
   \Gamma^{a}_{b c}::Depends(\partial{#}).

   R_{a b c d}::Depends(\nabla{#}).
   R^{a}_{b c d}::Depends(\nabla{#}).

\end{cadabra}

\clearpage

% =================================================================================================
\section*{Stage 1: Symmetrised partial derivatives of $g^{ab}$}

\begin{cadabra}
   beg_stage_1 = time.time()

   # symmetrised partial derivatives of g^{ab}

   gab00:=g^{a b}.                                              # cdb (gab00.101,gab00)

   gab01:= - g^{c b}\Gamma^{a}_{c d} - g^{a c}\Gamma^{b}_{c d}. # cdb (gab01.101,gab01)

   gab02:=\partial_{e}{ @(gab01) }.                             # cdb (gab02.101,gab02)
   distribute   (gab02)                                         # cdb (gab02.102,gab02)
   product_rule (gab02)                                         # cdb (gab02.103,gab02)
   substitute   (gab02, $\partial_{d}{g^{a b}} -> @(gab01)$)    # cdb (gab02.104,gab02)
   distribute   (gab02)                                         # cdb (gab02.105,gab02)

   gab03:=\partial_{f}{ @(gab02) }.                             # cdb (gab03.101,gab03)
   distribute   (gab03)                                         # cdb (gab03.102,gab03)
   product_rule (gab03)                                         # cdb (gab03.103,gab03)
   substitute   (gab03, $\partial_{d}{g^{a b}} -> @(gab01)$)    # cdb (gab03.104,gab03)
   distribute   (gab03)                                         # cdb (gab03.105,gab03)

   gab04:=\partial_{g}{ @(gab03) }.                             # cdb (gab04.101,gab04)
   distribute   (gab04)                                         # cdb (gab04.102,gab04)
   product_rule (gab04)                                         # cdb (gab04.103,gab04)
   substitute   (gab04, $\partial_{d}{g^{a b}} -> @(gab01)$)    # cdb (gab04.104,gab04)
   distribute   (gab04)                                         # cdb (gab04.105,gab04)

   gab05:=\partial_{h}{ @(gab04) }.                             # cdb (gab05.101,gab05)
   distribute   (gab05)                                         # cdb (gab05.102,gab05)
   product_rule (gab05)                                         # cdb (gab05.103,gab05)
   substitute   (gab05, $\partial_{d}{g^{a b}} -> @(gab01)$)    # cdb (gab05.104,gab05)
   distribute   (gab05)                                         # cdb (gab05.105,gab05)

   gab00 = impose_rnc (gab00)   # cdb (gab00.102,gab00)
   gab01 = impose_rnc (gab01)   # cdb (gab01.102,gab01)
   gab02 = impose_rnc (gab02)   # cdb (gab02.106,gab02)
   gab03 = impose_rnc (gab03)   # cdb (gab03.106,gab03)
   gab04 = impose_rnc (gab04)   # cdb (gab04.106,gab04)
   gab05 = impose_rnc (gab05)   # cdb (gab05.106,gab05)

\end{cadabra}

\clearpage

\begin{dgroup*}
   \begin{dmath*} \cdb*{gab00.101} \end{dmath*}
   \begin{dmath*} \cdb*{gab00.102} \end{dmath*}
   \begin{dmath*} \cdb*{gab01.101} \end{dmath*}
   \begin{dmath*} \cdb*{gab01.102} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{gab02.101} \end{dmath*}
   \begin{dmath*} \cdb*{gab02.102} \end{dmath*}
   \begin{dmath*} \cdb*{gab02.103} \end{dmath*}
   \begin{dmath*} \cdb*{gab02.104} \end{dmath*}
   \begin{dmath*} \cdb*{gab02.105} \end{dmath*}
   \begin{dmath*} \cdb*{gab02.106} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{gab03.101} \end{dmath*}
   \begin{dmath*} \cdb*{gab03.102} \end{dmath*}
   \begin{dmath*} \cdb*{gab03.103} \end{dmath*}
   \begin{dmath*} \cdb*{gab03.104} \end{dmath*}
   \begin{dmath*} \cdb*{gab03.105} \end{dmath*}
   \begin{dmath*} \cdb*{gab03.106} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{gab04.101} \end{dmath*}
   \begin{dmath*} \cdb*{gab04.102} \end{dmath*}
   \begin{dmath*} \cdb*{gab04.103} \end{dmath*}
   \begin{dmath*} \cdb*{gab04.104} \end{dmath*}
   \begin{dmath*} \cdb*{gab04.105} \end{dmath*}
   \begin{dmath*} \cdb*{gab04.106} \end{dmath*}
\end{dgroup*}

\begin{cadabra}
   # prepare first six terms in the Taylor series expansion of g^{ab}(x)

   term0:= @(gab00).
   distribute (term0)                             # cdb(term0.200,term0)

   term1:= @(gab01) A^d.
   distribute (term1)                             # cdb(term1.200,term1)

   term2:= @(gab02) A^d A^e.
   distribute (term2)                             # cdb(term2.200,term2)

   term3:= @(gab03) A^d A^e A^f.
   distribute (term3)                             # cdb(term3.200,term3)

   term4:= @(gab04) A^d A^e A^f A^g.
   distribute (term4)                             # cdb(term4.200,term4)

   term5:= @(gab05) A^d A^e A^f A^g A^h.
   distribute (term5)                             # cdb(term5.200,term5)

   end_stage_1 = time.time()
\end{cadabra}

\begin{dgroup*}
   \begin{dmath*} \cdb*{term0.200} \end{dmath*}
   \begin{dmath*} \cdb*{term1.200} \end{dmath*}
   \begin{dmath*} \cdb*{term2.200} \end{dmath*}
   \begin{dmath*} \cdb*{term3.200} \end{dmath*}
   % \begin{dmath*} \cdb*{term4.200} \end{dmath*}
   % \begin{dmath*} \cdb*{term5.200} \end{dmath*}
\end{dgroup*}

\clearpage

% =================================================================================================
\section*{Stage 2: Replace derivatives of $\Gamma$ with partial derivs of $R$}

\begin{cadabra}
   import cdblib

   beg_stage_2 = time.time()

   dGamma01 = cdblib.get ('dGamma01','dGamma.json')  # cdb(dGamma01.300,dGamma01)
   dGamma02 = cdblib.get ('dGamma02','dGamma.json')  # cdb(dGamma02.300,dGamma02)
   dGamma03 = cdblib.get ('dGamma03','dGamma.json')  # cdb(dGamma03.300,dGamma03)
   dGamma04 = cdblib.get ('dGamma04','dGamma.json')  # cdb(dGamma04.300,dGamma04)
   dGamma05 = cdblib.get ('dGamma05','dGamma.json')  # cdb(dGamma05.300,dGamma05)

   # replace partial derivs of \Gamma with products and derivs of Riemann tensor

   substitute (term2,$\partial_{c}{\Gamma^{a}_{b d}}A^{c}A^{b} -> @(dGamma01)$,repeat=True)                       # cdb(term2.301,term2)
   substitute (term2,$\partial_{c}{\Gamma^{a}_{d b}}A^{c}A^{b} -> @(dGamma01)$,repeat=True)                       # cdb(term2.302,term2)
   distribute (term2)                                                                                             # cdb(term2.303,term2)

   substitute (term3,$\partial_{c e}{\Gamma^{a}_{d b}}A^{c}A^{b}A^{e} -> @(dGamma02)$,repeat=True)                # cdb(term3.301,term3)
   substitute (term3,$\partial_{c e}{\Gamma^{a}_{b d}}A^{c}A^{b}A^{e} -> @(dGamma02)$,repeat=True)                # cdb(term3.302,term3)
   substitute (term3,$\partial_{c}{\Gamma^{a}_{b d}}A^{c}A^{b} -> @(dGamma01)$,repeat=True)                       # cdb(term3.303,term3)
   substitute (term3,$\partial_{c}{\Gamma^{a}_{d b}}A^{c}A^{b} -> @(dGamma01)$,repeat=True)                       # cdb(term3.304,term3)
   distribute (term3)                                                                                             # cdb(term3.305,term3)

   substitute (term4,$\partial_{c e f}{\Gamma^{a}_{d b}}A^{c}A^{b}A^{e}A^{f} -> @(dGamma03)$,repeat=True)         # cdb(term4.301,term4)
   substitute (term4,$\partial_{c e f}{\Gamma^{a}_{b d}}A^{c}A^{b}A^{e}A^{f} -> @(dGamma03)$,repeat=True)         # cdb(term4.302,term4)
   substitute (term4,$\partial_{c e}{\Gamma^{a}_{d b}}A^{c}A^{b}A^{e} -> @(dGamma02)$,repeat=True)                # cdb(term4.303,term4)
   substitute (term4,$\partial_{c e}{\Gamma^{a}_{b d}}A^{c}A^{b}A^{e} -> @(dGamma02)$,repeat=True)                # cdb(term4.304,term4)
   substitute (term4,$\partial_{c}{\Gamma^{a}_{b d}}A^{c}A^{b} -> @(dGamma01)$,repeat=True)                       # cdb(term4.305,term4)
   substitute (term4,$\partial_{c}{\Gamma^{a}_{d b}}A^{c}A^{b} -> @(dGamma01)$,repeat=True)                       # cdb(term4.306,term4)
   distribute (term4)                                                                                             # cdb(term4.307,term4)

   substitute (term5,$\partial_{c e f g}{\Gamma^{a}_{d b}}A^{c}A^{b}A^{e}A^{f}A^{g} -> @(dGamma04)$,repeat=True)  # cdb(term5.301,term5)
   substitute (term5,$\partial_{c e f g}{\Gamma^{a}_{b d}}A^{c}A^{b}A^{e}A^{f}A^{g} -> @(dGamma04)$,repeat=True)  # cdb(term5.302,term5)
   substitute (term5,$\partial_{c e f}{\Gamma^{a}_{d b}}A^{c}A^{b}A^{e}A^{f} -> @(dGamma03)$,repeat=True)         # cdb(term5.303,term5)
   substitute (term5,$\partial_{c e f}{\Gamma^{a}_{b d}}A^{c}A^{b}A^{e}A^{f} -> @(dGamma03)$,repeat=True)         # cdb(term5.304,term5)
   substitute (term5,$\partial_{c e}{\Gamma^{a}_{d b}}A^{c}A^{b}A^{e} -> @(dGamma02)$,repeat=True)                # cdb(term5.305,term5)
   substitute (term5,$\partial_{c e}{\Gamma^{a}_{b d}}A^{c}A^{b}A^{e} -> @(dGamma02)$,repeat=True)                # cdb(term5.306,term5)
   substitute (term5,$\partial_{c}{\Gamma^{a}_{b d}}A^{c}A^{b} -> @(dGamma01)$,repeat=True)                       # cdb(term5.307,term5)
   substitute (term5,$\partial_{c}{\Gamma^{a}_{d b}}A^{c}A^{b} -> @(dGamma01)$,repeat=True)                       # cdb(term5.308,term5)
   distribute (term5)                                                                                             # cdb(term5.309,term5)

   # ------------------------------------------------------------------------------------
   # this block only produces formatted output, it is not part of the main computation
   # ------------------------------------------------------------------------------------

   # the metric in terms of partial derivatives of Rabcd

   metric:=@(term0)
         + (1/1) @(term1)
         + (1/2) @(term2)
         + (1/6) @(term3)
         + (1/24) @(term4)
         + (1/120) @(term5).  # cdb(metric.301,metric)

   substitute (metric,$A^{a} -> x^{a}$)  # cdb (metric.302,metric)

   # reformat and tidy up

   Xterm0 := @(term0).
   Xterm1 := (1/1) @(term1).     # zero
   Xterm2 := (1/2) @(term2).
   Xterm3 := (1/6) @(term3).
   Xterm4 := (1/24) @(term4).
   Xterm5 := (1/120) @(term5).

   substitute (Xterm0,$A^{a} -> x^{a}$)
   substitute (Xterm1,$A^{a} -> x^{a}$)
   substitute (Xterm2,$A^{a} -> x^{a}$)
   substitute (Xterm3,$A^{a} -> x^{a}$)
   substitute (Xterm4,$A^{a} -> x^{a}$)
   substitute (Xterm5,$A^{a} -> x^{a}$)

   # Manipulating these expressions is hampered by the presence of the partial derivative on Rabcd.
   # Thus we can't freely rasie/lower indices on the dRabcd terms. But we can do so on the first
   # derivatives (since these are evaluated at x=0 where the connection vanishes).

   substitute       (Xterm2,$g^{a b} R^{c}_{d e b} -> R^{c}_{d e}^{a}}$)  # cdb(Xterm2.301,Xterm2)
   substitute       (Xterm3,$g^{a b} R^{c}_{d e b} -> R^{c}_{d e}^{a}}$)  # cdb(Xterm3.301,Xterm3)
   substitute       (Xterm4,$g^{a b} R^{c}_{d e b} -> R^{c}_{d e}^{a}}$)  # cdb(Xterm4.301,Xterm4)
   substitute       (Xterm5,$g^{a b} R^{c}_{d e b} -> R^{c}_{d e}^{a}}$)  # cdb(Xterm5.301,Xterm5)

   substitute       (Xterm2,$g^{b a} R^{c}_{d e b} -> R^{c}_{d e}^{a}}$)  # cdb(Xterm2.302,Xterm2)
   substitute       (Xterm3,$g^{b a} R^{c}_{d e b} -> R^{c}_{d e}^{a}}$)  # cdb(Xterm3.302,Xterm3)
   substitute       (Xterm4,$g^{b a} R^{c}_{d e b} -> R^{c}_{d e}^{a}}$)  # cdb(Xterm4.302,Xterm4)
   substitute       (Xterm5,$g^{b a} R^{c}_{d e b} -> R^{c}_{d e}^{a}}$)  # cdb(Xterm5.302,Xterm5)

   substitute       (Xterm2,$g^{a b} \partial_{c}{R^{d}_{e f b}} -> \partial_{c}{R^{d}_{e f}^{a}$)  # cdb(Xterm2.303,Xterm2)
   substitute       (Xterm3,$g^{a b} \partial_{c}{R^{d}_{e f b}} -> \partial_{c}{R^{d}_{e f}^{a}$)  # cdb(Xterm3.303,Xterm3)
   substitute       (Xterm4,$g^{a b} \partial_{c}{R^{d}_{e f b}} -> \partial_{c}{R^{d}_{e f}^{a}$)  # cdb(Xterm4.303,Xterm4)
   substitute       (Xterm5,$g^{a b} \partial_{c}{R^{d}_{e f b}} -> \partial_{c}{R^{d}_{e f}^{a}$)  # cdb(Xterm5.303,Xterm5)

   substitute       (Xterm2,$g^{b a} \partial_{c}{R^{d}_{e f b}} -> \partial_{c}{R^{d}_{e f}^{a}$)  # cdb(Xterm2.304,Xterm2)
   substitute       (Xterm3,$g^{b a} \partial_{c}{R^{d}_{e f b}} -> \partial_{c}{R^{d}_{e f}^{a}$)  # cdb(Xterm3.304,Xterm3)
   substitute       (Xterm4,$g^{b a} \partial_{c}{R^{d}_{e f b}} -> \partial_{c}{R^{d}_{e f}^{a}$)  # cdb(Xterm4.304,Xterm4)
   substitute       (Xterm5,$g^{b a} \partial_{c}{R^{d}_{e f b}} -> \partial_{c}{R^{d}_{e f}^{a}$)  # cdb(Xterm5.304,Xterm5)

   sort_product     (Xterm2)  # cdb(Xterm2.305,Xterm2)
   sort_product     (Xterm3)  # cdb(Xterm3.305,Xterm3)
   sort_product     (Xterm4)  # cdb(Xterm4.305,Xterm4)
   sort_product     (Xterm5)  # cdb(Xterm5.305,Xterm5)

   rename_dummies   (Xterm2)  # cdb(Xterm2.306,Xterm2)
   rename_dummies   (Xterm3)  # cdb(Xterm3.306,Xterm3)
   rename_dummies   (Xterm4)  # cdb(Xterm4.306,Xterm4)
   rename_dummies   (Xterm5)  # cdb(Xterm5.306,Xterm5)

   canonicalise     (Xterm2)  # cdb(Xterm2.307,Xterm2)
   canonicalise     (Xterm3)  # cdb(Xterm3.307,Xterm3)
   canonicalise     (Xterm4)  # cdb(Xterm4.307,Xterm4)
   canonicalise     (Xterm5)  # cdb(Xterm5.307,Xterm5)

   # We can simplify Xterm2 and Xterm3 by careful juggling of the indices (swapping free indices on selected terms)

   tmp = add_tags (Xterm2,'\\mu')       # cdb (tmp.001,tmp)
   zoom (tmp, $\mu_{1} Q??$)            # cdb (tmp.002,tmp)
   substitute (tmp, $R^{b}_{c d}^{a} x^{c} x^{d} -> R^{a}_{c d}^{b} x^{c} x^{d}$)  # cdb (tmp.003,tmp)
   unzoom (tmp)
   Xterm2 = clear_tags (tmp,'\\mu')     # cdb (Xterm2.401,Xterm2)

   tmp = add_tags (Xterm3,'\\mu')       # cdb (tmp.011,tmp)
   zoom (tmp, $\mu_{1} Q??$)            # cdb (tmp.012,tmp)
   substitute (tmp, $\partial_{c}{R^{b}_{d e}^{a}} x^{c} x^{d} x^{e} -> \partial_{c}{R^{a}_{d e}^{b}} x^{c} x^{d} x^{e}$)  # cdb (tmp.013,tmp)
   unzoom (tmp)
   Xterm3 = clear_tags (tmp,'\\mu')     # cdb (Xterm3.401,Xterm3)

   Xterm0 = reformat_xterm (Xterm0,  1)    # cdb(Xterm0.308,Xterm0)
   Xterm2 = reformat_xterm (Xterm2,  3)    # cdb(Xterm2.308,Xterm2)
   Xterm3 = reformat_xterm (Xterm3,  6)    # cdb(Xterm3.308,Xterm3)
   Xterm4 = reformat_xterm (Xterm4,360)    # cdb(Xterm4.308,Xterm4)
   Xterm5 = reformat_xterm (Xterm5,360)    # cdb(Xterm5.308,Xterm5)

   # metric to 4th and 6th order terms in powers of x^a

   Metric3 := @(Xterm0) + @(Xterm2).                                      # cdb (Metric3.301,Metric3)
   Metric4 := @(Xterm0) + @(Xterm2) + @(Xterm3).                          # cdb (Metric4.301,Metric4)
   Metric5 := @(Xterm0) + @(Xterm2) + @(Xterm3) + @(Xterm4).              # cdb (Metric5.301,Metric5)
   Metric6 := @(Xterm0) + @(Xterm2) + @(Xterm3) + @(Xterm4) + @(Xterm5).  # cdb (Metric6.301,Metric6)

   # ------------------------------------------------------------------------------------
   # end of format block
   # ------------------------------------------------------------------------------------

   end_stage_2 = time.time()
\end{cadabra}

\clearpage

\begin{dgroup*}
   \begin{dmath*} \cdb*{term2.301} \end{dmath*}
   \begin{dmath*} \cdb*{term2.302} \end{dmath*}
   \begin{dmath*} \cdb*{term2.303} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{term3.301} \end{dmath*}
   \begin{dmath*} \cdb*{term3.302} \end{dmath*}
   \begin{dmath*} \cdb*{term3.303} \end{dmath*}
   \begin{dmath*} \cdb*{term3.304} \end{dmath*}
   \begin{dmath*} \cdb*{term3.305} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{term4.301} \end{dmath*}
   \begin{dmath*} \cdb*{term4.302} \end{dmath*}
   \begin{dmath*} \cdb*{term4.303} \end{dmath*}
   \begin{dmath*} \cdb*{term4.304} \end{dmath*}
   \begin{dmath*} \cdb*{term4.305} \end{dmath*}
   \begin{dmath*} \cdb*{term4.306} \end{dmath*}
   \begin{dmath*} \cdb*{term4.307} \end{dmath*}
\end{dgroup*}

\clearpage

\begin{dgroup*}
   \begin{dmath*} g^{ab}(x) = \cdb{Metric3.301} \end{dmath*}
   \begin{dmath*} g^{ab}(x) = \cdb{Metric4.301} \end{dmath*}
   \begin{dmath*} g^{ab}(x) = \cdb{Metric5.301} \end{dmath*}
   \begin{dmath*} g^{ab}(x) = \cdb{Metric6.301} \end{dmath*}
\end{dgroup*}

\clearpage

% =================================================================================================
\section*{Stage 3: Replace partial derivs of $R$ with covariant derivs of $R$}

\begin{cadabra}
   beg_stage_3 = time.time()

   # now convert partial derivs of Rabcd to covariant derivs

   dRabcd01 = cdblib.get ('dRabcd01','dRabcd.json')  # cdb(dRabcd01.400,dRabcd01)
   dRabcd02 = cdblib.get ('dRabcd02','dRabcd.json')  # cdb(dRabcd02.400,dRabcd02)
   dRabcd03 = cdblib.get ('dRabcd03','dRabcd.json')  # cdb(dRabcd03.400,dRabcd03)

   # term1 & term2 need no special care, just a bit of tidying

   eliminate_metric (term1)   # cdb(term1.401,term1)
   sort_product     (term1)   # cdb(term1.402,term1)
   rename_dummies   (term1)   # cdb(term1.403,term1)
   canonicalise     (term1)   # cdb(term1.404,term1)

   eliminate_metric (term2)   # cdb(term2.401,term2)
   sort_product     (term2)   # cdb(term2.402,term2)
   rename_dummies   (term2)   # cdb(term2.403,term2)
   canonicalise     (term2)   # cdb(term2.404,term2)

   # replace partial derivatives of Riemann tensor in term3, term4 etc. with covariant derivatives of Rabcd

   tmp01 := @(dRabcd01).      # cdb(tmp01.403,tmp01)
   tmp02 := @(dRabcd02).      # cdb(tmp02.403,tmp02)
   tmp03 := @(dRabcd03).      # cdb(tmp03.403,tmp03)

   substitute (term3,$A^{c}A^{d}A^{e}\partial_{e}{R^{a}_{c d b}} ->   @(tmp01)$,repeat=True)         # cdb(term3.401,term3)
   substitute (term3,$A^{c}A^{d}A^{e}\partial_{e}{R^{a}_{c b d}} -> - @(tmp01)$,repeat=True)         # cdb(term3.402,term3)
   distribute (term3)                                                                                # cdb(term3.403,term3)

   substitute (term4,$A^{c}A^{d}A^{e}A^{f}\partial_{e f}{R^{a}_{c d b}} ->   @(tmp02)$,repeat=True)  # cdb(term4.401,term4)
   substitute (term4,$A^{c}A^{d}A^{e}A^{f}\partial_{e f}{R^{a}_{c b d}} -> - @(tmp02)$,repeat=True)  # cdb(term4.402,term4)
   substitute (term4,$A^{c}A^{d}A^{e}\partial_{e}{R^{a}_{c d b}} ->   @(tmp01)$,repeat=True)         # cdb(term4.403,term4)
   substitute (term4,$A^{c}A^{d}A^{e}\partial_{e}{R^{a}_{c b d}} -> - @(tmp01)$,repeat=True)         # cdb(term4.404,term4)
   distribute (term4)                                                                                # cdb(term4.405,term4)

   substitute (term5,$A^{c}A^{d}A^{e}A^{f}A^{g}\partial_{e f g}{R^{a}_{c d b}} ->   @(tmp03)$,repeat=True)
   substitute (term5,$A^{c}A^{d}A^{e}A^{f}A^{g}\partial_{e f g}{R^{a}_{c b d}} -> - @(tmp03)$,repeat=True)
   substitute (term5,$A^{c}A^{d}A^{e}A^{f}\partial_{e f}{R^{a}_{c d b}} ->   @(tmp02)$,repeat=True)
   substitute (term5,$A^{c}A^{d}A^{e}A^{f}\partial_{e f}{R^{a}_{c b d}} -> - @(tmp02)$,repeat=True)
   substitute (term5,$A^{c}A^{d}A^{e}\partial_{e}{R^{a}_{c d b}} ->   @(tmp01)$,repeat=True)
   substitute (term5,$A^{c}A^{d}A^{e}\partial_{e}{R^{a}_{c b d}} -> - @(tmp01)$,repeat=True)
   distribute (term5)

   end_stage_3 = time.time()
\end{cadabra}

\begin{dgroup*}
   \begin{dmath*} \cdb*{tmp01.403} \end{dmath*}
   \begin{dmath*} \cdb*{tmp02.403} \end{dmath*}
   \begin{dmath*} \cdb*{tmp03.403} \end{dmath*}
\end{dgroup*}

\clearpage

\begin{dgroup*}
   \begin{dmath*} \cdb*{term1.401} \end{dmath*}
   \begin{dmath*} \cdb*{term1.402} \end{dmath*}
   \begin{dmath*} \cdb*{term1.403} \end{dmath*}
   \begin{dmath*} \cdb*{term1.404} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{term2.401} \end{dmath*}
   \begin{dmath*} \cdb*{term2.402} \end{dmath*}
   \begin{dmath*} \cdb*{term2.403} \end{dmath*}
   \begin{dmath*} \cdb*{term2.404} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{term3.401} \end{dmath*}
   \begin{dmath*} \cdb*{term3.402} \end{dmath*}
   \begin{dmath*} \cdb*{term3.403} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{term4.401} \end{dmath*}
   \begin{dmath*} \cdb*{term4.402} \end{dmath*}
   \begin{dmath*} \cdb*{term4.403} \end{dmath*}
   \begin{dmath*} \cdb*{term4.404} \end{dmath*}
   \begin{dmath*} \cdb*{term4.405} \end{dmath*}
\end{dgroup*}

\clearpage

% =================================================================================================
\section*{Stage 4: Build the Taylor series for $g_{ab}$, reformatting and output}

\begin{cadabra}
   beg_stage_4 = time.time()

   # final housekeeping

   # lower the ^{ab} indices to _{uv}

   tmp0 := g_{a u} g_{b v} @(term0).
   tmp1 := g_{a u} g_{b v} @(term1).
   tmp2 := g_{a u} g_{b v} @(term2).
   tmp3 := g_{a u} g_{b v} @(term3).
   tmp4 := g_{a u} g_{b v} @(term4).
   tmp5 := g_{a u} g_{b v} @(term5).

   distribute           (tmp1)  # cdb(tmp1.501,tmp1)
   eliminate_metric     (tmp1)  # cdb(tmp1.502,tmp1)
   eliminate_kronecker  (tmp1)  # cdb(tmp1.503,tmp1)
   tmp1 = flatten_Rabcd (tmp1)
   canonicalise         (tmp1)  # cdb(tmp1.506,tmp1)

   distribute           (tmp2)  # cdb(tmp2.501,tmp2)
   eliminate_metric     (tmp2)  # cdb(tmp2.502,tmp2)
   eliminate_kronecker  (tmp2)  # cdb(tmp2.503,tmp2)
   tmp2 = flatten_Rabcd (tmp2)
   canonicalise         (tmp2)  # cdb(tmp2.506,tmp2)

   distribute           (tmp3)  # cdb(tmp3.501,tmp3)
   eliminate_metric     (tmp3)  # cdb(tmp3.502,tmp3)
   eliminate_kronecker  (tmp3)  # cdb(tmp3.503,tmp3)
   tmp3 = flatten_Rabcd (tmp3)
   canonicalise         (tmp3)  # cdb(tmp3.506,tmp3)

   distribute           (tmp4)  # cdb(tmp4.501,tmp4)
   eliminate_metric     (tmp4)  # cdb(tmp4.502,tmp4)
   eliminate_kronecker  (tmp4)  # cdb(tmp4.503,tmp4)
   tmp4 = flatten_Rabcd (tmp4)
   canonicalise         (tmp4)  # cdb(tmp4.506,tmp4)

   distribute           (tmp5)  # cdb(tmp5.501,tmp5)
   eliminate_metric     (tmp5)  # cdb(tmp5.502,tmp5)
   eliminate_kronecker  (tmp5)  # cdb(tmp5.503,tmp5)
   tmp5 = flatten_Rabcd (tmp5)
   canonicalise         (tmp5)  # cdb(tmp5.506,tmp5)

   # this is out final answer

   # raise the _{uv} indices to ^{ab}

   metric:= g^{a u} g^{b v} (  @(tmp0)
                              + (1/1) @(tmp1)
                              + (1/2) @(tmp2)
                              + (1/6) @(tmp3)
                              + (1/24) @(tmp4)
                              + (1/120) @(tmp5) ).  # cdb(metric.500,metric)

   distribute             (metric)   # cdb(metric.501,metric)
   eliminate_metric       (metric)   # cdb(metric.502,metric)
   eliminate_kronecker    (metric)   # cdb(metric.503,metric)
   metric = flatten_Rabcd (metric)   # cdb(metric.504,metric)
   canonicalise           (metric)   # cdb(metric.505,metric)

   substitute          (metric,$g_{a b} g^{b c} -> g_{a}^{c}$)
   substitute          (metric,$g_{b a} g^{b c} -> g_{a}^{c}$)
   substitute          (metric,$g_{b a} g^{c b} -> g_{a}^{c}$)
   substitute          (metric,$g_{a b} g^{c b} -> g_{a}^{c}$)
   eliminate_kronecker (metric)   # cdb(metric.506,metric)
   canonicalise        (metric)   # cdb(metric.507,metric)

   substitute (metric,$A^{a} -> x^{a}$)  # cdb (metric.508,metric)

   cdblib.create ('metric-inv.json')

   cdblib.put ('g^ab',metric,'metric-inv.json')

   # extract the terms of the metric in powers of x

   term0 = get_xterm (metric,0)   # cdb(term0.501,term0)
   term1 = get_xterm (metric,1)   # cdb(term1.501,term1)
   term2 = get_xterm (metric,2)   # cdb(term2.501,term2)
   term3 = get_xterm (metric,3)   # cdb(term3.501,term3)
   term4 = get_xterm (metric,4)   # cdb(term4.501,term4)
   term5 = get_xterm (metric,5)   # cdb(term5.501,term5)

   cdblib.put ('g^ab_0',term0,'metric-inv.json')
   cdblib.put ('g^ab_1',term1,'metric-inv.json')
   cdblib.put ('g^ab_2',term2,'metric-inv.json')
   cdblib.put ('g^ab_3',term3,'metric-inv.json')
   cdblib.put ('g^ab_4',term4,'metric-inv.json')
   cdblib.put ('g^ab_5',term5,'metric-inv.json')

   # this version of "metric" is used only in the commentary at the start of this notebook

   metric4:=@(term0) + @(term1) + @(term2) + @(term3).  # cdb(metric4.501,metric4)

\end{cadabra}

\clearpage

\begin{dgroup*}
   \begin{dmath*} \cdb*{term0.501} \end{dmath*}
   \begin{dmath*} \cdb*{term1.501} \end{dmath*}
   \begin{dmath*} \cdb*{term2.501} \end{dmath*}
   \begin{dmath*} \cdb*{term3.501} \end{dmath*}
   \begin{dmath*} \cdb*{term4.501} \end{dmath*}
   \begin{dmath*} \cdb*{term5.501} \end{dmath*}
\end{dgroup*}

\clearpage

\begin{dgroup*}
   \begin{dmath*} \cdb*{tmp2.501} \end{dmath*}
   \begin{dmath*} \cdb*{tmp2.502} \end{dmath*}
   \begin{dmath*} \cdb*{tmp2.503} \end{dmath*}
   \begin{dmath*} \cdb*{tmp2.506} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{tmp3.501} \end{dmath*}
   \begin{dmath*} \cdb*{tmp3.502} \end{dmath*}
   \begin{dmath*} \cdb*{tmp3.503} \end{dmath*}
   \begin{dmath*} \cdb*{tmp3.506} \end{dmath*}
\end{dgroup*}

\clearpage

\begin{dgroup*}
   \begin{dmath*} \cdb*{tmp4.501} \end{dmath*}
   \begin{dmath*} \cdb*{tmp4.502} \end{dmath*}
   \begin{dmath*} \cdb*{tmp4.503} \end{dmath*}
   \begin{dmath*} \cdb*{tmp4.506} \end{dmath*}
\end{dgroup*}

\clearpage

\begin{dgroup*}
   \begin{dmath*} \cdb*{tmp5.506} \end{dmath*}
\end{dgroup*}

\clearpage

\begin{dgroup*}
   \begin{dmath*} \cdb*{metric.500} \end{dmath*}
   \begin{dmath*} \cdb*{metric.501} \end{dmath*}
   \begin{dmath*} \cdb*{metric.502} \end{dmath*}
   \begin{dmath*} \cdb*{metric.503} \end{dmath*}
   \begin{dmath*} \cdb*{metric.504} \end{dmath*}
   \begin{dmath*} \cdb*{metric.505} \end{dmath*}
   \begin{dmath*} \cdb*{metric.506} \end{dmath*}
   \begin{dmath*} \cdb*{metric.507} \end{dmath*}
   \begin{dmath*} \cdb*{metric.508} \end{dmath*}
\end{dgroup*}

% =================================================================================================
% the remaining code is just for pretty printing

\clearpage

\begin{cadabra}
   Rterm0 := @(term0).
   Rterm1 := @(term1).  # zero
   Rterm2 := @(term2).
   Rterm3 := @(term3).
   Rterm4 := @(term4).
   Rterm5 := @(term5).

   Rterm0 = reformat_xterm (Rterm0,  1)    # cdb(Rterm0.601,Rterm0)
   Rterm2 = reformat_xterm (Rterm2,  3)    # cdb(Rterm2.601,Rterm2)
   Rterm3 = reformat_xterm (Rterm3,  6)    # cdb(Rterm3.601,Rterm3)
   Rterm4 = reformat_xterm (Rterm4, 60)    # cdb(Rterm4.601,Rterm4)
   Rterm5 = reformat_xterm (Rterm5, 90)    # cdb(Rterm5.601,Rterm5)

   Metric := @(Rterm0) + @(Rterm2) + @(Rterm3) + @(Rterm4) + @(Rterm5).  # cdb (Metric.601,Metric)

   scaled0 = rescale_xterm (Rterm0,  1)    # cdb(scaled0.601,scaled0)
   scaled2 = rescale_xterm (Rterm2,  3)    # cdb(scaled2.601,scaled2)
   scaled3 = rescale_xterm (Rterm3,  6)    # cdb(scaled3.601,scaled3)
   scaled4 = rescale_xterm (Rterm4, 60)    # cdb(scaled4.601,scaled4)
   scaled5 = rescale_xterm (Rterm5, 90)    # cdb(scaled5.601,scaled5)

   end_stage_4 = time.time()
\end{cadabra}

\clearpage

% =================================================================================================
\section*{The inverse metric in Riemann normal coordinates}

\begin{dgroup*}
   \begin{dmath*} g^{a b}(x) = \cdb{Metric.601}+\BigO{\eps^6} \end{dmath*}
\end{dgroup*}

\clearpage

% =================================================================================================
\section*{Curvature expansion of the inverse metric}
\begin{align*}
     g^{a b}(x) =
     \ngab{0}^{a b}
   + \ngab{2}^{a b}
   + \ngab{3}^{a b}
   + \ngab{4}^{a b}
   + \ngab{5}^{a b}+\BigO{\eps^6}
\end{align*}
\begin{dgroup*}
   \begin{dmath*}    \ngab{0}^{a b} = \cdb{scaled0.601} \end{dmath*}
   \begin{dmath*}  3 \ngab{2}^{a b} = \cdb{scaled2.601} \end{dmath*}
   \begin{dmath*}  6 \ngab{3}^{a b} = \cdb{scaled3.601} \end{dmath*}
   \begin{dmath*} 60 \ngab{4}^{a b} = \cdb{scaled4.601} \end{dmath*}
   \begin{dmath*} 90 \ngab{5}^{a b} = \cdb{scaled5.601} \end{dmath*}
\end{dgroup*}

\clearpage

% =================================================================================================
% export selected objects, these will later be imported into a library
% these are the objects that will appear in the paper

\begin{cadabra}
   cdblib.create ('metric-inv.export')

   cdblib.put ('g^ab_3',Metric3,'metric-inv.export')  # R and \partial R
   cdblib.put ('g^ab_4',Metric4,'metric-inv.export')
   cdblib.put ('g^ab_5',Metric5,'metric-inv.export')
   cdblib.put ('g^ab_6',Metric6,'metric-inv.export')

   cdblib.put ('g^ab',  Metric, 'metric-inv.export')  # R and \nabla R

   cdblib.put ('g^ab_scaled0',scaled0,'metric-inv.export')
   cdblib.put ('g^ab_scaled2',scaled2,'metric-inv.export')
   cdblib.put ('g^ab_scaled3',scaled3,'metric-inv.export')
   cdblib.put ('g^ab_scaled4',scaled4,'metric-inv.export')
   cdblib.put ('g^ab_scaled5',scaled5,'metric-inv.export')

   checkpoint.append (Metric4)
   checkpoint.append (Metric6)

   checkpoint.append (Metric)

   checkpoint.append (scaled0)
   checkpoint.append (scaled2)
   checkpoint.append (scaled3)
   checkpoint.append (scaled4)
   checkpoint.append (scaled5)

   # cdbBeg (timing)
   print ("Stage 1: {:7.1f} secs\\hfill\\break".format(end_stage_1-beg_stage_1))
   print ("Stage 2: {:7.1f} secs\\hfill\\break".format(end_stage_2-beg_stage_2))
   print ("Stage 3: {:7.1f} secs\\hfill\\break".format(end_stage_3-beg_stage_3))
   print ("Stage 4: {:7.1f} secs".format(end_stage_4-beg_stage_4))
   # cdbEnd (timing)

\end{cadabra}

\clearpage

% =================================================================================================
\section*{Timing}

\cdb{timing}

% =================================================================================================
% export checkpoints in json format

\bgroup
\CdbSetup{action=hide}
\begin{cadabra}
   for i in range( len(checkpoint) ):
      cdblib.put ('check{:03d}'.format(i),checkpoint[i],checkpoint_file)
\end{cadabra}
\egroup

\end{document}