### 2019-09-24 source/cadabra/sqrt-detg.tex

\documentclass[12pt]{pylatex}

\begin{document}

Expand $\sqrt{g}$ in powers of $R,\nabla R$ etc.

\begin{python}
   from sympy import *

   eps, a, b, c, d = symbols('\epsilon a b c d')

   ans = sqrt(1+a*eps**2+b*eps**3+c*eps**4+d*eps**5)   # py(ans.001,ans)
   ans = ans.series(eps, 0, 6)                         # py(ans.002,ans)
   ans = simplify(ans)                                 # py(ans.003,ans)

\end{python}

\begin{dgroup*}
   \begin{dmath*} \py*{ans.001} \end{dmath*}
   \begin{dmath*} \py*{ans.002} \end{dmath*}
   \begin{dmath*} \py*{ans.003} \end{dmath*}
\end{dgroup*}

And while we're here, let's also expand $\log(g)$ in powers of $R,\nabla R$ etc.

\begin{python}
   from sympy import *

   eps, a, b, c, d = symbols('\epsilon a b c d')

   ans = log(1+a*eps**2+b*eps**3+c*eps**4+d*eps**5)    # py(ans.101,ans)
   ans = ans.series(eps, 0, 6)                         # py(ans.102,ans)
   ans = simplify(ans)                                 # py(ans.103,ans)

\end{python}

\begin{dgroup*}
   \begin{dmath*} \py*{ans.101} \end{dmath*}
   \begin{dmath*} \py*{ans.102} \end{dmath*}
   \begin{dmath*} \py*{ans.103} \end{dmath*}
\end{dgroup*}

\end{document}