### 2019-09-24 source/cadabra/tests/semantic/dGamma.tex

\documentclass[12pt]{cdblatex}

\begin{document}

\section*{\jobname}

\CdbSetup{action=hide}

\begin{cadabra}
   import shared

   import cdblib

   term00A = cdblib.get ('check000','expected/dGamma.json')
   term01A = cdblib.get ('check001','expected/dGamma.json')
   term02A = cdblib.get ('check002','expected/dGamma.json')
   term03A = cdblib.get ('check003','expected/dGamma.json')
   term04A = cdblib.get ('check004','expected/dGamma.json')

   term00B = cdblib.get ('check000','output/dGamma.json')
   term01B = cdblib.get ('check001','output/dGamma.json')
   term02B = cdblib.get ('check002','output/dGamma.json')
   term03B = cdblib.get ('check003','output/dGamma.json')
   term04B = cdblib.get ('check004','output/dGamma.json')

   diff000 = shared.check (term00A,term00B)   # cdb (diff000,diff000)
   diff001 = shared.check (term01A,term01B)   # cdb (diff001,diff001)
   diff002 = shared.check (term02A,term02B)   # cdb (diff002,diff002)
   diff003 = shared.check (term03A,term03B)   # cdb (diff003,diff003)
   diff004 = shared.check (term04A,term04B)   # cdb (diff004,diff004)

\end{cadabra}

\begin{dgroup*}
   \Dmath*{ \cdb*{diff000} }
   \Dmath*{ \cdb*{diff001} }
   \Dmath*{ \cdb*{diff002} }
   \Dmath*{ \cdb*{diff003} }
   \Dmath*{ \cdb*{diff004} }
\end{dgroup*}

\end{document}