### 2019-10-18 source/cadabra/sqrt-detg.tex
Updated Makefile to *not* make sqrt-detg
The sqrt-detg code requires pylatex.sh and friends which are not included in this repo. (see https://github.com/leo-brewin/hybrid-latex). The sqrt-detg results are not required by any code here. It's output was hand coded in detg.tex.

Also, the cdbpreproc.py was updated (copied from hybrid-latex repo.).
See https://github.com/leo-brewin/hybrid-latex/commit/199d8e817b54ca515c1221aee749990829b69538

\documentclass[12pt]{pylatex}

\begin{document}

Expand $\sqrt{g}$ in powers of $R,\nabla R$ etc.

\begin{python}
   from sympy import *

   eps, a, b, c, d = symbols('\epsilon a b c d')

   ans = sqrt(1+a*eps**2+b*eps**3+c*eps**4+d*eps**5)   # py (ans.001,ans)
   ans = ans.series(eps, 0, 6)                         # py (ans.002,ans)
   ans = simplify(ans)                                 # py (ans.003,ans)

\end{python}

\begin{dgroup*}
   \begin{dmath*} \py*{ans.001} \end{dmath*}
   \begin{dmath*} \py*{ans.002} \end{dmath*}
   \begin{dmath*} \py*{ans.003} \end{dmath*}
\end{dgroup*}

And while we're here, let's also expand $\log(g)$ in powers of $R,\nabla R$ etc.

\begin{python}
   from sympy import *

   eps, a, b, c, d = symbols('\epsilon a b c d')

   ans = log(1+a*eps**2+b*eps**3+c*eps**4+d*eps**5)    # py (ans.101,ans)
   ans = ans.series(eps, 0, 6)                         # py (ans.102,ans)
   ans = simplify(ans)                                 # py (ans.103,ans)

\end{python}

\begin{dgroup*}
   \begin{dmath*} \py*{ans.101} \end{dmath*}
   \begin{dmath*} \py*{ans.102} \end{dmath*}
   \begin{dmath*} \py*{ans.103} \end{dmath*}
\end{dgroup*}

\end{document}