### 2020-02-20 source/cadabra/geodesic-ivp.tex
Corrected +/- sign error on s^2, s^3 etc.  terms in eqs. 11.19a-d
I had forgotten to take account of the minus sign in eq. 3.1 when constructing the solution to the geodesic IVP. This correction has no effect on any other results.

\documentclass[12pt]{cdblatex}

\lstset{gobble=2}

\begin{document}

% =================================================================================================
% create checkpoint file

\bgroup
\CdbSetup{action=hide}
\begin{cadabra}
   import cdblib
   checkpoint_file = 'tests/semantic/output/geodesic-ivp.json'
   cdblib.create (checkpoint_file)
   checkpoint = []
\end{cadabra}
\egroup

% =================================================================================================
\section*{Geodesic IVP}

Our game here is to find the solution of
\begin{align*}
   0 = \frac{d^2 x^{a}}{ds^2} + \Gamma^{a}_{bc}(x) \frac{dx^b}{ds} \frac{dx^c}{ds}
\end{align*}
subject to the initial conditions $x^{a}(s) = x^a$ and $dx^a(s)/ds={\Dot x}^{a}$ at $s=0$.

% =================================================================================================
\section*{Algorithm}

By successive differentiation of the above equation we can compute
\begin{align*}
   \frac{d^n x^{a}}{ds^n} = -\Gamma^{a}_{\udn}\frac{dx^{\udn}}{ds}
\end{align*}
at $s=0$ for $n=2,3,4,\dotsc$. The $\Gamma^{a}_{\udn}$ are the \emph{generalised connections}.

We can then construct the Taylor series solution for $x^{a}(s)$
\begin{align*}
   x^a(s) = x^a + s {\Dot x}^a - \sum_{k=2}^\infty\>\frac{s^{k}}{k!} \Gamma^{a}_{\udk}{\Dot x}^{\udk}
\end{align*}

\clearpage

\begin{cadabra}
   {a,b,c,d,e,f,g,h,i,j,k,l,m,n,o,p,q,r,s,t,u,v,w#}::Indices(position=independent).

   \nabla{#}::Derivative.

   import cdblib

   # change signs to account for - sign in front of the sum for x^a(s), see above preamble

   def flip_sign (obj):
       return Ex(0) - obj

   sterm21 = flip_sign (cdblib.get ('genGamma01','genGamma.json'))
   sterm22 = flip_sign (cdblib.get ('genGamma02','genGamma.json'))
   sterm23 = flip_sign (cdblib.get ('genGamma03','genGamma.json'))
   sterm24 = flip_sign (cdblib.get ('genGamma04','genGamma.json'))

   sterm31 = flip_sign (cdblib.get ('genGamma11','genGamma.json'))
   sterm32 = flip_sign (cdblib.get ('genGamma12','genGamma.json'))
   sterm33 = flip_sign (cdblib.get ('genGamma13','genGamma.json'))

   sterm41 = flip_sign (cdblib.get ('genGamma21','genGamma.json'))
   sterm42 = flip_sign (cdblib.get ('genGamma22','genGamma.json'))

   sterm51 = flip_sign (cdblib.get ('genGamma31','genGamma.json'))

   sterm2 := @(sterm21) + @(sterm22) + @(sterm23) + @(sterm24).  # cdb (sterm2.000,sterm2)
   sterm3 := @(sterm31) + @(sterm32) + @(sterm33).               # cdb (sterm3.000,sterm3)
   sterm4 := @(sterm41) + @(sterm42).                            # cdb (sterm4.000,sterm4)
   sterm5 := @(sterm51).                                         # cdb (sterm5.000,sterm5)

   factor_out (sterm2,$A^{a?}$)                                  # cdb (sterm2.001,sterm2)
   factor_out (sterm3,$A^{a?}$)                                  # cdb (sterm3.001,sterm3)
   factor_out (sterm4,$A^{a?}$)                                  # cdb (sterm4.001,sterm4)
   factor_out (sterm5,$A^{a?}$)                                  # cdb (sterm5.001,sterm5)

   sterm2 := 360 @(sterm2).
   sterm3 := 360 @(sterm3).
   sterm4 :=  90 @(sterm4).
   sterm5 :=   3 @(sterm5).

   substitute (sterm2,$A^{a}->1$)                                # cdb (sterm2.002,sterm2)
   substitute (sterm3,$A^{a}->1$)                                # cdb (sterm3.002,sterm3)
   substitute (sterm4,$A^{a}->1$)                                # cdb (sterm4.002,sterm4)
   substitute (sterm5,$A^{a}->1$)                                # cdb (sterm5.002,sterm5)

\end{cadabra}

% =================================================================================================
% the remaining code is just for pretty printing

\clearpage

% =================================================================================================
\section*{The geodesic ivp}

\begin{align*}
   x^{a}(s) = x^{a}
            + s {\dot{x}^a}
            + \frac{s^2}{2!} {\dot{x}^b} {\dot{x}^c} A^{a}_{bc}
            + \frac{s^3}{3!} {\dot{x}^b} {\dot{x}^c} {\dot{x}^d} A^{a}_{bcd}
            + \frac{s^4}{4!} {\dot{x}^b} {\dot{x}^c} {\dot{x}^d} {\dot{x}^e} A^{a}_{bcde}
            + \frac{s^5}{5!} {\dot{x}^b} {\dot{x}^c} {\dot{x}^d} {\dot{x}^e} {\dot{x}^f} A^{a}_{bcdef}
            + \dotsb
\end{align*}
\begin{dgroup*}
   \begin{dmath*} 360 A^{a}_{bc} = \cdb{sterm2.002} \end{dmath*}
   \begin{dmath*} 360 A^{a}_{bcd} = \cdb{sterm3.002} \end{dmath*}
   \begin{dmath*}  90 A^{a}_{bcde} = \cdb{sterm4.002} \end{dmath*}
   \begin{dmath*}   3 A^{a}_{bcdef} = \cdb{sterm5.002} \end{dmath*}
\end{dgroup*}

\clearpage

% =================================================================================================
% export selected objects, these will later be imported into a library
% these are the objects that will appear in the paper

\begin{cadabra}
   sterm2short := @(sterm21) + @(sterm22).             # cdb (sterm2.short.001,sterm2short)
   sterm3short := @(sterm31).                          # cdb (sterm3.short.001,sterm3short)
   sterm2shortscaled := 12 @(sterm2short).             # cdb (sterm2.short.scaled.002,sterm2shortscaled)
   sterm3shortscaled :=  2 @(sterm3short).             # cdb (sterm3.short.scaled.002,sterm3shortscaled)

   substitute (sterm2shortscaled,$A^{a}->1$)           # cdb (sterm2.short.scaled.003,sterm2shortscaled)
   substitute (sterm3shortscaled,$A^{a}->1$)           # cdb (sterm3.short.scaled.003,sterm3shortscaled)

   cdblib.create ('geodesic-ivp.export')

   # 4th order ivp terms scaled
   cdblib.put ('ivp42',sterm2shortscaled,'geodesic-ivp.export')
   cdblib.put ('ivp43',sterm3shortscaled,'geodesic-ivp.export')

   # 6th order ivp terms scaled
   cdblib.put ('ivp62',sterm2,'geodesic-ivp.export')
   cdblib.put ('ivp63',sterm3,'geodesic-ivp.export')
   cdblib.put ('ivp64',sterm4,'geodesic-ivp.export')
   cdblib.put ('ivp65',sterm5,'geodesic-ivp.export')

   checkpoint.append (sterm2shortscaled)
   checkpoint.append (sterm3shortscaled)

   checkpoint.append (sterm2)
   checkpoint.append (sterm3)
   checkpoint.append (sterm4)
   checkpoint.append (sterm5)
\end{cadabra}

% just to check that we are exporting the correct 4th order terms

\begin{dgroup*}
   \begin{dmath*} \cdb*{sterm2.short.001} \end{dmath*}
   \begin{dmath*} \cdb*{sterm3.short.001} \end{dmath*}
   \begin{dmath*} \cdb*{sterm2.short.scaled.002} \end{dmath*}
   \begin{dmath*} \cdb*{sterm3.short.scaled.002} \end{dmath*}
   \begin{dmath*} \cdb*{sterm2.short.scaled.003} \end{dmath*}
   \begin{dmath*} \cdb*{sterm3.short.scaled.003} \end{dmath*}
\end{dgroup*}

% =================================================================================================
% export checkpoints in json format

\bgroup
\CdbSetup{action=hide}
\begin{cadabra}
   for i in range( len(checkpoint) ):
      cdblib.put ('check{:03d}'.format(i),checkpoint[i],checkpoint_file)
\end{cadabra}
\egroup

\end{document}