### 2024-10-28 source/cadabra/geodesic-ivp.tex
Extended output to 2nd, 3rd and 4th order
Extended the Cadabra output to include 2d, 3rd, 4th and 5th order expressions for the metric, inverse metric and the geodesic-ivp. These can be used in other Cadabra code for other computations.

\def\Date{30 Jul 2024}
% \def\FileID{file:}

\documentclass[12pt]{cdblatex}

\lstset{gobble=2}

\begin{document}

% =================================================================================================
% create checkpoint file

\bgroup
\CdbSetup{action=hide}
\begin{cadabra}
   import cdblib
   checkpoint_file = 'tests/semantic/output/geodesic-ivp.json'
   cdblib.create (checkpoint_file)
   checkpoint = []
\end{cadabra}
\egroup

% =================================================================================================
\section*{Geodesic IVP}

Our game here is to find the solution of
\begin{align*}
   0 = \frac{d^2 x^{a}}{ds^2} + \Gamma^{a}_{bc}(x) \frac{dx^b}{ds} \frac{dx^c}{ds}
\end{align*}
subject to the initial conditions $x^{a}(s) = x^a$ and $dx^a(s)/ds={\Dot x}^{a}$ at $s=0$.

% =================================================================================================
\section*{Algorithm}

By successive differentiation of the above equation we can compute
\begin{align*}
   \frac{d^n x^{a}}{ds^n} = -\Gamma^{a}_{\udn}\frac{dx^{\udn}}{ds}
\end{align*}
at $s=0$ for $n=2,3,4,\dotsc$. The $\Gamma^{a}_{\udn}$ are the \emph{generalised connections}.

We can then construct the Taylor series solution for $x^{a}(s)$
\begin{align*}
   x^a(s) = x^a + s {\Dot x}^a - \sum_{k=2}^\infty\>\frac{s^{k}}{k!} \Gamma^{a}_{\udk}{\Dot x}^{\udk}
\end{align*}

\clearpage

\begin{cadabra}
   {a,b,c,d,e,f,g,h,i,j,k,l,m,n,o,p,q,r,s,t,u,v,w#}::Indices(position=independent).

   \nabla{#}::Derivative.

   import cdblib

   # change signs to account for - sign in front of the sum for x^a(s), see above preamble

   def flip_sign (obj):
       return Ex(0) - obj

   sterm21 = flip_sign (cdblib.get ('genGamma01','genGamma.json'))
   sterm22 = flip_sign (cdblib.get ('genGamma02','genGamma.json'))
   sterm23 = flip_sign (cdblib.get ('genGamma03','genGamma.json'))
   sterm24 = flip_sign (cdblib.get ('genGamma04','genGamma.json'))

   sterm31 = flip_sign (cdblib.get ('genGamma11','genGamma.json'))
   sterm32 = flip_sign (cdblib.get ('genGamma12','genGamma.json'))
   sterm33 = flip_sign (cdblib.get ('genGamma13','genGamma.json'))

   sterm41 = flip_sign (cdblib.get ('genGamma21','genGamma.json'))
   sterm42 = flip_sign (cdblib.get ('genGamma22','genGamma.json'))

   sterm51 = flip_sign (cdblib.get ('genGamma31','genGamma.json'))

   # note: the various ivp21, ivp31  etc. are the pieces of the Taylor series
   #       for the ivp but *without* the leading 1/n! of the Taylor series

   ivp21 := @(sterm21).                                          # cdb (ivp21.000,ivp21)

   ivp31 := @(sterm21) + @(sterm22).                             # cdb (ivp31.000,ivp31)
   ivp32 := @(sterm31).                                          # cdb (ivp32.000,ivp32)

   ivp41 := @(sterm21) + @(sterm22) + @(sterm23).                # cdb (ivp41.000,ivp41)
   ivp42 := @(sterm31) + @(sterm32).                             # cdb (ivp42.000,ivp42)
   ivp43 := @(sterm41).                                          # cdb (ivp43.000,ivp43)

   ivp51 := @(sterm21) + @(sterm22) + @(sterm23) + @(sterm24).   # cdb (ivp51.000,ivp51)
   ivp52 := @(sterm31) + @(sterm32) + @(sterm33).                # cdb (ivp52.000,ivp52)
   ivp53 := @(sterm41) + @(sterm42).                             # cdb (ivp53.000,ivp53)
   ivp54 := @(sterm51).                                          # cdb (ivp54.000,ivp54)

   factor_out (ivp21,$A^{a?}$)                                   # cdb (ivp21.001,ivp21)

   factor_out (ivp31,$A^{a?}$)                                   # cdb (ivp31.001,ivp31)
   factor_out (ivp32,$A^{a?}$)                                   # cdb (ivp32.001,ivp32)

   factor_out (ivp41,$A^{a?}$)                                   # cdb (ivp41.001,ivp41)
   factor_out (ivp42,$A^{a?}$)                                   # cdb (ivp42.001,ivp42)
   factor_out (ivp43,$A^{a?}$)                                   # cdb (ivp43.001,ivp43)

   factor_out (ivp51,$A^{a?}$)                                   # cdb (ivp51.001,ivp51)
   factor_out (ivp52,$A^{a?}$)                                   # cdb (ivp52.001,ivp52)
   factor_out (ivp53,$A^{a?}$)                                   # cdb (ivp53.001,ivp53)
   factor_out (ivp54,$A^{a?}$)                                   # cdb (ivp54.001,ivp54)

   v{#}::LaTeXForm("\dot{x}").

   substitute (ivp21, $A^{a} -> v^{a}$)                          # cdb (ivp21.002,ivp21)

   substitute (ivp31, $A^{a} -> v^{a}$)                          # cdb (ivp31.002,ivp31)
   substitute (ivp32, $A^{a} -> v^{a}$)                          # cdb (ivp32.002,ivp32)

   substitute (ivp41, $A^{a} -> v^{a}$)                          # cdb (ivp41.002,ivp41)
   substitute (ivp42, $A^{a} -> v^{a}$)                          # cdb (ivp42.002,ivp42)
   substitute (ivp43, $A^{a} -> v^{a}$)                          # cdb (ivp43.002,ivp43)

   substitute (ivp51, $A^{a} -> v^{a}$)                          # cdb (ivp51.002,ivp51)
   substitute (ivp52, $A^{a} -> v^{a}$)                          # cdb (ivp52.002,ivp52)
   substitute (ivp53, $A^{a} -> v^{a}$)                          # cdb (ivp53.002,ivp53)
   substitute (ivp54, $A^{a} -> v^{a}$)                          # cdb (ivp54.002,ivp54)

   # build the Taylor series
   # note the inclusion of the 1/n! factors

   ivp2 := x^{a} + s v^{a} + (1/2) (s**2) @(ivp21).
   ivp3 := x^{a} + s v^{a} + (1/2) (s**2) @(ivp31) + (1/6) (s**3) @(ivp32).
   ivp4 := x^{a} + s v^{a} + (1/2) (s**2) @(ivp41) + (1/6) (s**3) @(ivp42) + (1/24) (s**4) @(ivp43).
   ivp5 := x^{a} + s v^{a} + (1/2) (s**2) @(ivp51) + (1/6) (s**3) @(ivp52) + (1/24) (s**4) @(ivp53) + (1/120) (s**5) @(ivp54).

   # cdb (ivp2.000,ivp2)
   # cdb (ivp3.000,ivp3)
   # cdb (ivp4.000,ivp4)
   # cdb (ivp5.000,ivp5)

   # now construct the scaled terms for ivp5

   sterm2 := @(sterm21) + @(sterm22) + @(sterm23) + @(sterm24).  # cdb (sterm2.000,sterm2)
   sterm3 := @(sterm31) + @(sterm32) + @(sterm33).               # cdb (sterm3.000,sterm3)
   sterm4 := @(sterm41) + @(sterm42).                            # cdb (sterm4.000,sterm4)
   sterm5 := @(sterm51).                                         # cdb (sterm5.000,sterm5)

   factor_out (sterm2,$A^{a?}$)                                  # cdb (sterm2.001,sterm2)
   factor_out (sterm3,$A^{a?}$)                                  # cdb (sterm3.001,sterm3)
   factor_out (sterm4,$A^{a?}$)                                  # cdb (sterm4.001,sterm4)
   factor_out (sterm5,$A^{a?}$)                                  # cdb (sterm5.001,sterm5)

   sterm2 := 360 @(sterm2).
   sterm3 := 360 @(sterm3).
   sterm4 :=  90 @(sterm4).
   sterm5 :=   3 @(sterm5).

   substitute (sterm2,$A^{a}->1$)                                # cdb (sterm2.002,sterm2)
   substitute (sterm3,$A^{a}->1$)                                # cdb (sterm3.002,sterm3)
   substitute (sterm4,$A^{a}->1$)                                # cdb (sterm4.002,sterm4)
   substitute (sterm5,$A^{a}->1$)                                # cdb (sterm5.002,sterm5)

\end{cadabra}

% =================================================================================================
% the remaining code is just for pretty printing

\clearpage

% =================================================================================================
\section*{The geodesic ivp}

\begin{align*}
   x^{a}(s) = x^{a}
            + s {\dot{x}^a}
            + \frac{s^2}{2!} {\dot{x}^b} {\dot{x}^c} A^{a}_{bc}
            + \frac{s^3}{3!} {\dot{x}^b} {\dot{x}^c} {\dot{x}^d} A^{a}_{bcd}
            + \frac{s^4}{4!} {\dot{x}^b} {\dot{x}^c} {\dot{x}^d} {\dot{x}^e} A^{a}_{bcde}
            + \frac{s^5}{5!} {\dot{x}^b} {\dot{x}^c} {\dot{x}^d} {\dot{x}^e} {\dot{x}^f} A^{a}_{bcdef}
            + \dotsb
\end{align*}
\begin{dgroup*}
   \begin{dmath*} 360 A^{a}_{bc} = \cdb{sterm2.002} \end{dmath*}
   \begin{dmath*} 360 A^{a}_{bcd} = \cdb{sterm3.002} \end{dmath*}
   \begin{dmath*}  90 A^{a}_{bcde} = \cdb{sterm4.002} \end{dmath*}
   \begin{dmath*}   3 A^{a}_{bcdef} = \cdb{sterm5.002} \end{dmath*}
\end{dgroup*}

\clearpage

% =================================================================================================
% export selected objects, these will later be imported into a library
% these are the objects that will appear in the paper

\begin{cadabra}
   sterm2short := @(sterm21) + @(sterm22).             # cdb (sterm2.short.001,sterm2short)
   sterm3short := @(sterm31).                          # cdb (sterm3.short.001,sterm3short)
   sterm2shortscaled := 12 @(sterm2short).             # cdb (sterm2.short.scaled.002,sterm2shortscaled)
   sterm3shortscaled :=  2 @(sterm3short).             # cdb (sterm3.short.scaled.002,sterm3shortscaled)

   substitute (sterm2shortscaled,$A^{a}->1$)           # cdb (sterm2.short.scaled.003,sterm2shortscaled)
   substitute (sterm3shortscaled,$A^{a}->1$)           # cdb (sterm3.short.scaled.003,sterm3shortscaled)

   cdblib.create ('geodesic-ivp.export')

   # 4th order ivp terms scaled
   cdblib.put ('ivp42',sterm2shortscaled,'geodesic-ivp.export')
   cdblib.put ('ivp43',sterm3shortscaled,'geodesic-ivp.export')

   # 6th order ivp terms scaled
   cdblib.put ('ivp62',sterm2,'geodesic-ivp.export')
   cdblib.put ('ivp63',sterm3,'geodesic-ivp.export')
   cdblib.put ('ivp64',sterm4,'geodesic-ivp.export')
   cdblib.put ('ivp65',sterm5,'geodesic-ivp.export')

   checkpoint.append (sterm2shortscaled)
   checkpoint.append (sterm3shortscaled)

   checkpoint.append (sterm2)
   checkpoint.append (sterm3)
   checkpoint.append (sterm4)
   checkpoint.append (sterm5)

   cdblib.create ('geodesic-ivp.json')

   cdblib.put ('ivp21',ivp21,'geodesic-ivp.json')

   cdblib.put ('ivp31',ivp31,'geodesic-ivp.json')
   cdblib.put ('ivp32',ivp32,'geodesic-ivp.json')

   cdblib.put ('ivp41',ivp41,'geodesic-ivp.json')
   cdblib.put ('ivp42',ivp42,'geodesic-ivp.json')
   cdblib.put ('ivp43',ivp43,'geodesic-ivp.json')

   cdblib.put ('ivp51',ivp51,'geodesic-ivp.json')
   cdblib.put ('ivp52',ivp52,'geodesic-ivp.json')
   cdblib.put ('ivp53',ivp53,'geodesic-ivp.json')
   cdblib.put ('ivp54',ivp54,'geodesic-ivp.json')

   cdblib.put ('ivp2',ivp2,'geodesic-ivp.json')
   cdblib.put ('ivp3',ivp3,'geodesic-ivp.json')
   cdblib.put ('ivp4',ivp4,'geodesic-ivp.json')
   cdblib.put ('ivp5',ivp5,'geodesic-ivp.json')
\end{cadabra}

% just to check that we are exporting the correct 4th order terms

\begin{dgroup*}
   \begin{dmath*} \cdb*{sterm2.short.001} \end{dmath*}
   \begin{dmath*} \cdb*{sterm3.short.001} \end{dmath*}
   \begin{dmath*} \cdb*{sterm2.short.scaled.002} \end{dmath*}
   \begin{dmath*} \cdb*{sterm3.short.scaled.002} \end{dmath*}
   \begin{dmath*} \cdb*{sterm2.short.scaled.003} \end{dmath*}
   \begin{dmath*} \cdb*{sterm3.short.scaled.003} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{ivp21.002} \end{dmath*}
   \begin{dmath*} \cdb*{ivp31.002} \end{dmath*}
   \begin{dmath*} \cdb*{ivp32.002} \end{dmath*}
   \begin{dmath*} \cdb*{ivp41.002} \end{dmath*}
   \begin{dmath*} \cdb*{ivp42.002} \end{dmath*}
   \begin{dmath*} \cdb*{ivp43.002} \end{dmath*}
   \begin{dmath*} \cdb*{ivp51.002} \end{dmath*}
   \begin{dmath*} \cdb*{ivp52.002} \end{dmath*}
   \begin{dmath*} \cdb*{ivp53.002} \end{dmath*}
   \begin{dmath*} \cdb*{ivp54.002} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{ivp2.000} \end{dmath*}
   \begin{dmath*} \cdb*{ivp3.000} \end{dmath*}
   \begin{dmath*} \cdb*{ivp4.000} \end{dmath*}
   % \begin{dmath*} \cdb*{ivp5.000} \end{dmath*}
\end{dgroup*}

% =================================================================================================
% export checkpoints in json format

\bgroup
\CdbSetup{action=hide}
\begin{cadabra}
   for i in range( len(checkpoint) ):
      cdblib.put ('check{:03d}'.format(i),checkpoint[i],checkpoint_file)
\end{cadabra}
\egroup

\end{document}