### 2025-04-12 source/cadabra/sqrt-detg.tex
Removed dependancy on cadabra2python
I used cadabra2python as a backdoor way to import .cdb files.
This would create a .py file which would then be imported using
a standard Python import statement.

But Cadabra can now (since at least version 2.5.x) import .cdb files directly.
So I no longer need the backdoor hack.

\def\Date{19 Jan 2024}
% \def\FileID{file:}

\documentclass[12pt]{pylatex}
\usepackage{config}

\begin{document}

% =================================================================================================
\section*{Side calculations for detg2.tex}

This is a short computation for the expansions of $\sqrt{-\det g(x)}$ and $\log\left(-\det g(x)\right)$
in powers of $R$ and its derivatives. The results wil be used in {\tts detg2.tex}.

The starting point is to write $\det g(x)$ in the form
\begin{dgroup*}
   \begin{dmath*} -g(x) = 1 + a \epsilon^2 + b \epsilon^3 + c \epsilon^4 + d \epsilon^5 + \BigO{\epsilon^6} \end{dmath*}
\end{dgroup*}
where $\eps$ is introduced as an expansion parameter and where $a,b,c$ and $d$
are simple expressions in $R$ and its derivatives. These can be read off directly
from the expansion for $\det g(x)$ given in {\tts detg2.tex}.

\clearpage

\begin{python}
   from sympy import *

   eps, a, b, c, d = symbols('\epsilon a b c d')

   ans = sqrt(1+a*eps**2+b*eps**3+c*eps**4+d*eps**5)   # py (ans.001,ans)
   ans = ans.series(eps, 0, 6)                         # py (ans.002,ans)
   ans = simplify(ans)                                 # py (ans.003,ans)

   det = 1 + eps**2*a + eps**3*b + eps**4*c + eps**5*d   # py (det.001,det)
   # this foo will be used in detg2.tex
   foo = Rational(1,2) + det/2 - (eps**4*a**2)/8 - (eps**5*a*b)/4   # py (foo.001,foo)
   err = simplify (ans-foo)   # py (err.001,err)

\end{python}

\begin{dgroup*}
   \begin{dmath*} \py*{ans.001} \end{dmath*}
   \begin{dmath*} \py*{ans.002} \end{dmath*}
   \begin{dmath*} \py*{ans.003} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \py*{det.001} \end{dmath*}
   \begin{dmath*} \py*{foo.001} \end{dmath*}
   \begin{dmath*} \py*{err.001} \end{dmath*}
\end{dgroup*}

\clearpage

And while we're here, let's also expand $\log(g)$ in powers of $R,\nabla R$ etc.

\begin{python}
   from sympy import *

   eps, a, b, c, d = symbols('\epsilon a b c d')

   ans = log(1+a*eps**2+b*eps**3+c*eps**4+d*eps**5)    # py (ans.101,ans)
   ans = ans.series(eps, 0, 6)                         # py (ans.102,ans)
   ans = simplify(ans)                                 # py (ans.103,ans)

   det = 1 + eps**2*a + eps**3*b + eps**4*c + eps**5*d   # py (det.001,det)
   # this foo will be used in detg2.tex
   foo = -1 + det - (eps**4*a**2)/2 - (eps**5*a*b)       # py (foo.002,foo)
   err = simplify (ans-foo)   # py (err.002,err)
\end{python}

\begin{dgroup*}
   \begin{dmath*} \py*{ans.101} \end{dmath*}
   \begin{dmath*} \py*{ans.102} \end{dmath*}
   \begin{dmath*} \py*{ans.103} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \py*{det.001} \end{dmath*}
   \begin{dmath*} \py*{foo.002} \end{dmath*}
   \begin{dmath*} \py*{err.002} \end{dmath*}
\end{dgroup*}

\end{document}